### 2022-10-03 Manuscript/Paper.tex
First Commit

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage[margin=1 in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

\title{STAT3494}
\author{Colin Sullivan\\
University of Connecticut}
\date{October 3rd, 2022}

\begin{document}
\maketitle

\begin{abstract}

This will contain the abstract for my paper.

\end{abstract}

\section{Introduction}

What is the topic and why is it worth studying? - the first major section of text in the paper, the Introduction commonly describes the topic under investigation, summarizes or discusses relevant prior research (for related details, please see the Writing Literature Reviews section of this website), identifies unresolved issues that the current research will address, and provides an overview of the research that is to be described in greater detail in the sections to follow.

To cite a reference, here are examples.
\citet{xie2015dynamic} did something ...

A lot of work has been done \citep[e.g.,][]{xie2015dynamic}.

\section{Data}

\begin{equation}
  \label{eq:mc2}
  E= m c^2
\end{equation}

\section{Methods}

What did you do? - a section which details how the research was performed.  It typically features a description of the participants/subjects that were involved, the study design, the materials that were used, and the study procedure.  If there were multiple experiments, then each experiment may require a separate Methods section.  A rule of thumb is that the Methods section should be sufficiently detailed for another researcher to duplicate your research.

\begin{equation}
	\label{eq:area}
	\pi r^2
\end{equation}

Equation~\eqref{eq:area} is the area of a circle. Below is an unnumbered equation.

\[
f(x)=\frac{1}{\sqrt{3\pi}}\exp
\]

\section{Results}

What did you find? - a section which describes the data that was collected and the results of any statistical tests that were performed.  It may also be prefaced by a description of the analysis procedure that was used. If there were multiple experiments, then each experiment may require a separate Results section.

Table~\ref{tab:rv} summarizes some example draws from some distributions.

\begin{table}[ht]
  \caption{This is a table of some distributions.}
  \label{tab:rv}
\centering
\begin{tabular}{rrr}
  \toprule
normal & poisson & gamma \\ 
  \midrule
-0.110 & 4 & 2.401 \\ 
  0.116 & 4 & 3.529 \\ 
  -0.828 & 9 & 2.112 \\ 
  -0.066 & 6 & 11.104 \\ 
  0.219 & 3 & 4.815 \\ 
  0.303 & 5 & 2.188 \\ 
  0.544 & 0 & 8.050 \\ 
  -2.617 & 8 & 3.646 \\ 
  0.747 & 1 & 5.178 \\ 
  -1.103 & 4 & 3.043 \\ 
   \bottomrule
\end{tabular}
\end{table}

Figure~\ref{fig:cars} shows the distance against the speed from this dataset.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{cars.pdf}
  \caption{This is a figure of the cars.pdf}
  \label{fig:cars}
\end{figure}

\section{Discussion}
What is the significance of your results? - the final major section of text in the paper.  The Discussion commonly features a summary of the results that were obtained in the study, describes how those results address the topic under investigation and/or the issues that the research was designed to address, and may expand upon the implications of those findings.  Limitations and directions for future research are also commonly addressed.

Another citation \citep{srinath2017python}.

\bibliography{references}
\bibliographystyle{chicago}

\end{document}

### 2022-10-05 Proposal/Proposal.tex
Update Proposal.tex

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% for space filling
\usepackage{lipsum}
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data

\title{Proposal: WNBA vs NBA Earnings In Comparable Players}
\author{Colin Sullivan\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\paragraph{Introduction}
I have chosen to take a close examination of WNBA and NBA players statistics and salaries. There is not a huge amount of current work on the subject, and it was difficult to find a solid enough data set of WNBA players and salaries. This is an important and engaging subject to me because I love basketball and want to see the game grow as much as possible, and hopefully with some statistical analysis of the data I can find trends into why the WNBA salaries are so exorbitantly lower than the NBA's for equivalent players.

\paragraph{Specific Aims}
I would like to compare not just salaries of WNBA vs NBA players, which is something that has been done numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, and statistical performances for each player. I am interested to learn if a similarly performing player in the WNBA gets a comparable salary (adjusted for league revenue/average salaries) to an NBA player, if there are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, and similar questions.

\paragraph{Data}
I have two data sets that I have been manipulating using a python script I wrote to make them more readable. The first is a collection of WNBA players in the 2018 league season and a number of descriptive statistics, including their age, number of games played, and various basketball stats. I also amended it to include their salaries using a separate website link. The second data set is similar but for NBA players in the same season. I plan to amend both to include league average salaries and league income to be able to make league adjusted statistics.

\paragraph{Research Design and Methods}
I plan to conduct numerous hypothesis tests on the questions that I outlined in the introduction section, using p values and t-tests to compare average statistics and salaries across leagues. These would be independent t tests. I also plan on utilizing bootstrapping to conduct hundreds or thousands of such tests using Python to make sure that I am getting the most accurate results not affected by strange outliers (there are a few). I am considering incorporating some regression in to predict future stats, but I am unsure if that will be feasible with my current data sets.

\paragraph{Discussion}
I expect to find that the NBA players simply get more money for similar performances as that of WNBA players. I also expect to find that they receive more adjusted compensation, although that remains to be seen and is the basis of the comparison idea. I hope that my work will further the conversation on gender equal pay across sports, an issue as widespread as any and not limited to only basketball. I also hope that my work will provide concrete data results to drive future changes should the leagues attempt to lesser the gap between player compensations. If the investigation goes differently then expected, it will be a very pleasant surprise (although may not bode well for the efforts of female players to increase pay).

\paragraph{Conclusion}
To summarize: My goal in this paper is to further the discussion on the pay gap between gender in basketball and how this holds back the game as a whole. I hope to do this by manipulating existing data sets to include new stats of my own design, which I will then compare to identical stats in the opposite league using hypothesis tests, t-tests, and bootstrapping (and potentially other tests) to come to conclusions regarding the issue and hopefully be pointed in the right direction to find a solution.

\bibliography{../manuscript/refs}
\bibliographystyle{chicago}

\end{document}

### 2022-10-11 Proposal/Proposal.tex
Update Proposal.tex
Remove lipsum, cannot be compiled on my computer
Update Proposal.tex
Updated proposal refs
Commit
Updated Proposals with full citations

@@
 % for space filling
-\usepackage{lipsum}
 % highlighting hyper links
@@
 \paragraph{Introduction}
-I have chosen to take a close examination of WNBA and NBA players statistics and salaries.
-There is not a huge amount of current work on the subject, and it was difficult to find a solid enough data set of WNBA players and salaries.
-This is an important and engaging subject to me because I love basketball and want to see the game grow as much as possible, and hopefully with some statistical analysis of the data I can find trends into why the WNBA salaries are so exorbitantly lower than the NBA's for equivalent players.
+I have chosen to take a close examination of WNBA and NBA players statistics and salaries. It was 
+extremely difficult to find a solid enough data set of WNBA players and salaries. This is an important 
+and engaging subject to me because I love basketball and want to see the game grow as much as possible, 
+and hopefully with some statistical analysis of the data I can find trends into why the WNBA salaries 
+are so exorbitantly lower than the NBA's for equivalent players.
+\newline
+There have been numerous studies surrounding that of the NBA and the WNBA and the differences in salaries.
+For example, one such paper authored by Elle Baker discusses the intra-league salary distribution and
+how WNBA players, despite getting paid far less, have much less "salary inequality"
+\cite{baker2020comparison}. Another argues for an increase in salaries in the WNBA (a cause everyone
+should be able to get behind) by citing the league's growth in fans and revenues \cite{ettienne2019s}.
+There seems to be a lack of specific adjusted salary comparisons, however, and I hope to build off articles
+like these to further the conversation.
 
 \paragraph{Specific Aims}
-I would like to compare not just salaries of WNBA vs NBA players, which is something that has been done numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, and statistical performances for each player.
-I am interested to learn if a similarly performing player in the WNBA gets a comparable salary (adjusted for league revenue/average salaries) to an NBA player, if there are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, and similar questions.
+I would like to compare not just salaries of WNBA vs NBA players, which is something that has been 
+done numerous times, but to go further in depth and examine salaries vs team salary caps, league 
+revenues, games played, and statistical performances for each player. I am interested to learn i
+f a similarly performing player in the WNBA gets a comparable salary (adjusted for league revenue/average
+salaries) to an NBA player, if there are certain statistics in the NBA that result in higher adjusted 
+salaries than the WNBA and vice versa, and similar questions.
+These are similar, yet more in depth questions to the ones posed by Elle Baker.
 
 \paragraph{Data}
-I have two data sets that I have been manipulating using a python script I wrote to make them more readable.
-The first is a collection of WNBA players in the 2018 league season and a number of descriptive statistics, including their age, number of games played, and various basketball stats.
-I also amended it to include their salaries using a separate website link.
-The second data set is similar but for NBA players in the same season.
-I plan to amend both to include league average salaries and league income to be able to make league adjusted statistics.
+I have two data sets that I have been manipulating using a python script I wrote to make them more 
+readable. The first is a collection of WNBA players \cite{first} in the 2018 league season and a number of 
+descriptive statistics, including their age, number of games played, and various basketball 
+stats. I also amended it to include their salaries using a separate website link. The second data 
+set is similar but for NBA players in the same season \cite{nba}. I plan to amend both to include league average 
+salaries and league income to be able to make league adjusted statistics.
 
 \paragraph{Research Design and Methods}
-I plan to conduct numerous hypothesis tests on the questions that I outlined in the introduction section, using p values and t-tests to compare average statistics and salaries across leagues.
-These would be independent t tests.
-I also plan on utilizing bootstrapping to conduct hundreds or thousands of such tests using Python to make sure that I am getting the most accurate results not affected by strange outliers (there are a few).
-I am considering incorporating some regression in to predict future stats, but I am unsure if that will be feasible with my current data sets.
+I plan to conduct numerous hypothesis tests on the questions that I outlined in the introduction 
+section, using p values and t-tests to compare average statistics and salaries across leagues. 
+These would be independent t tests. I also plan on utilizing bootstrapping to conduct hundreds 
+or thousands of such tests using Python to make sure that I am getting the most accurate results 
+not affected by strange outliers (there are a few). I am considering incorporating some regression 
+in to predict future stats, but I am unsure if that will be feasible with my current data sets.
 
 \paragraph{Discussion}
-I expect to find that the NBA players simply get more money for similar performances as that of WNBA players.
-I also expect to find that they receive more adjusted compensation, although that remains to be seen and is the basis of the comparison idea.
-I hope that my work will further the conversation on gender equal pay across sports, an issue as widespread as any and not limited to only basketball.
-I also hope that my work will provide concrete data results to drive future changes should the leagues attempt to lesser the gap between player compensations.
-If the investigation goes differently then expected, it will be a very pleasant surprise (although may not bode well for the efforts of female players to increase pay).
+I expect to find that the NBA players simply get more money for similar performances as that of WNBA 
+players. I also expect to find that they receive more adjusted compensation, although that remains to 
+be seen and is the basis of the comparison idea. I hope that my work will further the conversation on 
+gender equal pay across sports, an issue as widespread as any and not limited to only basketball. 
+I also hope that my work will provide concrete data results to drive future changes should the leagues 
+attempt to lesser the gap between player compensations. If the investigation goes differently then 
+expected, it will be a very pleasant surprise (although may not bode well for the efforts of female 
+players to increase pay).
 
 \paragraph{Conclusion}
-To summarize: My goal in this paper is to further the discussion on the pay gap between gender in basketball and how this holds back the game as a whole.
-I hope to do this by manipulating existing data sets to include new stats of my own design, which I will then compare to identical stats in the opposite league using hypothesis tests, t-tests, and bootstrapping (and potentially other tests) to come to conclusions regarding the issue and hopefully be pointed in the right direction to find a solution.
+To summarize: My goal in this paper is to further the discussion on the pay gap between gender 
+in basketball and how this holds back the game as a whole. I hope to do this by manipulating 
+existing data sets to include new stats of my own design, which I will then compare to identical 
+stats in the opposite league using hypothesis tests, t-tests, and bootstrapping (and potentially 
+other tests) to come to conclusions regarding the issue and hopefully be pointed in the right 
+direction to find a solution.
 
-\bibliography{../manuscript/refs}
+\bibliography{../Proposal/refs}
 \bibliographystyle{chicago}

### 2022-11-05 Code/WNBA_Cleanup.py
WNBA Cleanup Python Script
v1.0, need to write it to the csv and upload new csvs

@@ -1,0 +1,46 @@
+import csv
+
+## player_ID,Player,year_ID,Age,Tm,tm_gms,Tm_Net_Rtg,Pos,G,MP,MP_pct,PER,TS_pct,ThrPAr,FTr,ORB_pct,TRB_pct,AST_pct,STL_pct,BLK_pct,TOV_pct,USG_pct,OWS,DWS,WS,WS40,Composite_Rating,Wins_Generated
+
+lst = []
+with open("wnba-player-stats.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        lst.append(row)
+
+playerstat = []
+for player in lst:
+    if player[2] == "2018":
+        playerstat.append(player)
+# print(lst2)
+
+
+
+lst_sal = []
+with open("WNBA_Salaries.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        lst_sal.append(row)
+# print(lst_sal)
+del lst_sal[0]
+lst_salaries = []
+for player in lst_sal:
+    newlist = []
+    player[1] = player[1].replace('\xa0', ' ')
+    newlist.append(player[1])
+    salary = int(player[2])
+    if salary > 0:
+        newlist.append(player[2])
+    lst_salaries.append(newlist)
+# print(lst_salaries)
+# print(playerstat[0])
+
+for player in playerstat:
+    for salary in lst_salaries:
+        if salary[0] == player[1]:
+#             print(salary)
+            if len(salary) > 1:
+                player[0] = salary[1]
+#             print(player)
+print(playerstat)
+    

### 2022-11-05 Code/WNBA_Cleanup.py
Updated to write to new csv file

@@ -44,3 +44,14 @@ for player in playerstat:
 #             print(player)
 print(playerstat)
     
+    
+header = ['Player_Salary','Player','year_ID','Age','Tm','tm_gms','Tm_Net_Rtg','Pos','G','MP','MP_pct','PER',
+          'TS_pct','ThrPAr','FTr','ORB_pct','TRB_pct','AST_pct','STL_pct','BLK_pct','TOV_pct','USG_pct','OWS',
+          'DWS','WS','WS40','Composite_Rating','Wins_Generated']
+print(header)
+
+with open('WNBA_Combined.csv', 'w') as f:
+    writer = csv.writer(f)
+    writer.writerow(header)
+    for row in playerstat:
+        writer.writerow(row)

### 2022-11-05 Code/NBA_Cleanup.py
NBA Data Cleanup
My NBA data cleanup python file, combining data from two datasets and cleaning up, removing duplicates, etc.

@@ -1,0 +1,65 @@
+import csv
+
+## Player,Pos,Age,Tm,G,MP,PER,TS%,3PAr,FTr,ORB%,DRB%,TRB%,AST%,STL%,BLK%,TOV%,USG%,OWS,DWS,WS,WS/48,OBPM,DBPM,BPM,VORP
+
+lst = []
+with open("nba.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        lst.append(row)
+
+# players_2018 = []
+# for player in lst:
+#     if player[2] == "2019":
+#         players_2018.append(player)
+# # print(players_2018)
+
+for row in lst:
+    row[1] = row[1].split('\\')[0]
+#     print(row[1])
+salary_lst = []
+with open("nba-salaries.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        if row[-1] == "2019":
+            salary_lst.append(row)
+# print(salary_lst)
+
+apended_list = []
+for player in lst:
+    for salary in salary_lst:
+        if player[1] == salary[1]:
+            player[0] = salary[-2]
+            apended_list.append(player)
+# print(apended_list)
+# my_str = "Alex Abrines\abrinal01"
+# result = my_str.split('\\')
+# print(result)
+
+i = 0
+while i < len(apended_list):
+    if apended_list[i][4] == "TOT":
+        name = apended_list[i][1]
+        j = 0
+        while j < 4:
+            if apended_list[i+j][1] == name:
+                del apended_list[i+j]
+            j += 1
+    i += 1
+            
+print(apended_list)
+
+header = ['Salary','Player','Pos','Age','Tm','G','MP','PER','TS%','3PAr','FTr','ORB%','DRB%','TRB%','AST%',
+          'STL%','BLK%','TOV%','USG%','OWS','DWS','WS','WS/48','OBPM','DBPM','BPM','VORP']
+
+          
+          
+with open('NBA_Combined.csv', 'w') as f:
+    writer = csv.writer(f)
+    writer.writerow(header)
+    for row in apended_list:
+        writer.writerow(row)
+
+# header = ['Player_Salary','Player','year_ID','Age','Tm','tm_gms','Tm_Net_Rtg','Pos','G','MP','MP_pct','PER',
[…]

### 2022-11-05 Code/NBA_Cleanup.py
Update NBA_Cleanup.py
Added function to remove outliers (under 10 games played)

@@ -52,7 +52,13 @@ print(apended_list)
 header = ['Salary','Player','Pos','Age','Tm','G','MP','PER','TS%','3PAr','FTr','ORB%','DRB%','TRB%','AST%',
           'STL%','BLK%','TOV%','USG%','OWS','DWS','WS','WS/48','OBPM','DBPM','BPM','VORP']
 
+for player in apended_list:
+    while("" in player):
+        player.remove("")
           
+for player in apended_list:
+    if int(player[5]) < 10:
+        apended_list.remove(player)     
           
 with open('NBA_Combined.csv', 'w') as f:
     writer = csv.writer(f)

### 2022-11-07 Manuscript/Proposal.tex
Added Intro and most of data section

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% for space filling
% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

%% meta data
\setlength\parindent{24pt}
\linespread{2}
\title{WNBA vs NBA Earnings In Comparable Players}
\author{Colin Sullivan\\
  Department of Statistics\\
  University of Connecticut
}

\begin{document}
\maketitle

\paragraph{Introduction}
Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates among fans comparing the female counterpart to the much older, more popular National Basketball Association (NBA). Women’s basketball has a rich history marred by frequent sexism, fewer opportunities, and significantly less financial support.
\par
Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented \cite{Shattering_The_Glass}. Due to the gender roles and cultural norms, however, women at the time were believed to be very fragile and unable to play physical sports like basketball at anywhere near the same level as men and thus played under modified rules specially created for them \cite{WNBA_Hist}. Those familiar with basketball will know that it is usually played competitively with 5 players on each team, and each player runs the full court to play offense on one end and defend the other team on the other end. In 1892, however, women initially played “9 on 9”, with the court divided into three areas: offense, defense, and midcourt. Each woman was assigned one of these courts in which they played, such that 3 women on each team were in each section. The players could not leave their section of the court to go into the other, essentially creating three separate “3 on 3” games. The other major difference from the men’s game was that women could only dribble the ball a limited number of times and could not hold it for more than three seconds, rule changed attributed to the fear that the women players might suffer “nervous fatigue” if the game were too strenuous \cite{WNBA_Hist}.
\par
Everyone now knows that such attitudes towards women’s sports are ridiculous and sexist. Certainly there are some physical characteristic differences between men and women in sports, but for every male athlete there is a just as accomplished female one. The gradual changing of perspectives on women’s athletics is seen in the evolution of women’s basketball; while men’s basketball has largely stayed the same over the course of its almost 130 year history (key changes including the legalization of the slam dunk and the implementation of the three point line), women’s basketball has shifted slowly to eventually become almost identical. The sport was played inter-collegiately for decades beginning in the late 19th century, but did not enter the professional realm until the 1970s. The sport became officially recognized at the collegiate level by the NCAA in 1982 (a shockingly late introduction), and the Women’s Basketball League (WBL), founded in 1978 and only lasting three seasons, was the first of several failed women’s professional leagues in the late 20th century. In the meantime, the NBA experienced a huge boom in popularity largely due to the emergence of superstars Larry Bird, Magic Johnson, and Michael Jordan. The massive increase in NBA reach and revenue directly led the league creating, funding, and supporting the WNBA in 1996.
\par
There are minimal differences between today's women’s basketball and the basketball that men have been playing since inception. The WNBA and NCAAW ball is slightly smaller than that used by male players, the three point line is slightly shorter, and the game clock features four 10 minute quarters as opposed to the NBA’s 12 minute quarters, but for all intents and purposes the games are the same. The rules are otherwise identical, the athletes are similarly talented, the level of play is incredibly high, and the games are very enjoyable to watch. There remains only one massive difference between the two leagues: player salaries.
\par
The difference in player salaries is a huge point of contention in women’s basketball discourse and is often the subject of intense ridicule, and it is easy to see why: top NBA players make upwards of 40 million USD per season, not including endorsement deals and other income, while the WNBA most valuable player makes less than the average office worker in New York (an exaggeration, maybe, but Elena Delle Donne made around 120 thousand USD in her MVP year with little extra endorsement money or exposure). An NBA rookie signs one contract and has enough wealth to live comfortably off for the rest of his life, often for generations; a WNBA prospect is encouraged to continue her academic studies to ensure that she is able to get a good job after her career is over. Now, granted, the league revenue is much much less than that of the NBA, but many questions still remain that this paper hopes to address: Are the players paid fairly compared to league revenue? Are they paid fairly compared to NBA players based on basketball production? How can this change?
\par
There have been numerous studies surrounding that of the NBA and the WNBA and the differences in salaries.
For example, one such paper authored by Elle Baker discusses the intra-league salary distribution and
how WNBA players, despite getting paid far less, have much less "salary inequality"
\cite{baker2020comparison}. Another argues for an increase in salaries in the WNBA (a cause everyone
should be able to get behind) by citing the league's growth in fans and revenues \cite{ettienne2019s}.
There seems to be a lack of specific adjusted salary comparisons, however, and I hope to build off articles
like these to further the conversation.
\par
In this paper, I hope to compare not just salaries of WNBA vs NBA players, which is something that has been done numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, and statistical performances for each player. The goal is to draw conclusions based on the analysis to learn if a similarly performing player in the WNBA gets a comparable salary (adjusted for league revenue/average salaries) to an NBA player, if there are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, and similar questions. These are related, yet more in depth questions to the ones posed by Elle Baker.
\par
The paper will begin by discussing the datasets being analyzed and how they were cleaned and prepped for analysis using Python scripts. In the next section, statistical analysis will be performed to try and answer some of the questions posed in this introduction. Following that will be a discussion on the results and hypotheses for best paths forward. Finally, there will be a conclusions section to wrap up the paper and review the results.

\paragraph{Data}
There are two primary data sets that I will be using to conduct the analysis in this paper. The first of these is a database of historical WNBA data dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed on his GitHub account \cite{first}. Unlike most data sets, this one contains a multitude of advanced statistics for each player. The statistical categories are as follows:
\newline
\par
Player Name, Year, Age, Team, Team Games, Team Net Rating, Position, Games, Minutes Played, Minutes Played (as a percentage of team total minutes), Player Efficiency Rating, True Shooting \%, Free Throw Rating, Offensive Rebounding \%, Total Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, Win Shares per 40 games, Composite Rating, and Wins Generated.
\newline
\par
Using another data set from Paine’s Github, a Python script called WNBA\_Cleanup.py was written and used to combine salary data from the two sources and filter out the desired candidate players from the data set. This paper will focus on comparing players from the 2018 NBA and WNBA seasons, so that filter was applied. The script also filtered out players who played less than 5 games and those that did not have salary data, so as to make the set as standardized as possible and remove outliers. The final data set after filtration and combination was called WNBA\_Combined.csv and has data (the aforementioned statistical categories) for 166 WNBA players from the 2018 season.
\par
A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site Kaggle \cite{nba}. This data set contains a list of all players from the NBA in the 2017-18 season, with similar advanced statistics to the WNBA dataset:
\newline
\par
Player Name, Position, Age, Team, Games Played, Minutes Played, Player Efficiency Rating, True Shooting \%, 3 Point Attempt Rate, Free Throw Rate, Offensive Rebounding \%, Defensive Rebounding \%, Total Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, Win Shares per 48 games, Offensive Box +/-, Defensive Box +/-, Box +/-, and VORP.
\newline
\par
The data set notably did not contain player salaries, so a second NBA data set was necessary. One was found on GitHub, created by Erik Gregory Webb, which contains the player salaries for every NBA player from 2000-2020 \cite{nba_salaries}. In a similar manner to the WNBA data, a Python script called NBA\_Cleanup.py was created to combine the data from the two datasets into one csv file. The data was then cleaned up by removing outliers (players playing under 5 games) and those that did not have salary data. The script was also used to combine data for players who played on multiple teams; the original data set would have players listed separately for each team they played on. The final data set after filtration and combinated was called NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) for 376 NBA players from the 2017-18 season.

\paragraph{Research Design and Methods}
I plan to conduct numerous hypothesis tests on the questions that I outlined in the introduction 
section, using p values and t-tests to compare average statistics and salaries across leagues. 
These would be independent t tests. I also plan on utilizing bootstrapping to conduct hundreds 
or thousands of such tests using Python to make sure that I am getting the most accurate results 
not affected by strange outliers (there are a few). I am considering incorporating some regression 
in to predict future stats, but I am unsure if that will be feasible with my current data sets.

\paragraph{Discussion}
I expect to find that the NBA players simply get more money for similar performances as that of WNBA 
players. I also expect to find that they receive more adjusted compensation, although that remains to 
be seen and is the basis of the comparison idea. I hope that my work will further the conversation on 
gender equal pay across sports, an issue as widespread as any and not limited to only basketball. 
I also hope that my work will provide concrete data results to drive future changes should the leagues 
attempt to lesser the gap between player compensations. If the investigation goes differently then 
expected, it will be a very pleasant surprise (although may not bode well for the efforts of female 
players to increase pay).

\paragraph{Conclusion}
To summarize: My goal in this paper is to further the discussion on the pay gap between gender 
in basketball and how this holds back the game as a whole. I hope to do this by manipulating 
existing data sets to include new stats of my own design, which I will then compare to identical 
stats in the opposite league using hypothesis tests, t-tests, and bootstrapping (and potentially 
other tests) to come to conclusions regarding the issue and hopefully be pointed in the right 
direction to find a solution.

\bibliography{../Proposal/refs}
\bibliographystyle{chicago}

\end{document}

### 2022-11-09 Code/LR_Test_NBA.py
Uploading code that runs linear regression, NBA

@@ -1,0 +1,87 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+from scipy import stats
+
+lst = []
+with open("NBA_Combined.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        lst.append(row)
+
+
+
+# print(x)
+# print(y)
+
+def estimate_coef(x, y):
+	# number of observations/points
+	n = np.size(x)
+
+	# mean of x and y vector
+	m_x = np.mean(x)
+	m_y = np.mean(y)
+
+	# calculating cross-deviation and deviation about x
+	SS_xy = np.sum(y*x) - n*m_y*m_x
+	SS_xx = np.sum(x*x) - n*m_x*m_x
+
+	# calculating regression coefficients
+	b_1 = SS_xy / SS_xx
+	b_0 = m_y - b_1*m_x
+
+	return (b_0, b_1)
+
+def plot_regression_line(x, y, b, stat):
+	# plotting the actual points as scatter plot
+	plt.scatter(x, y, color = "m",
+			marker = "o", s = 30)
+
+	# predicted response vector
+	y_pred = b[0] + b[1]*x
+
+	# plotting the regression line
+	plt.plot(x, y_pred, color = "g")
+
+	# putting labels
+	plt.title(stat)
+	plt.xlabel('x')
+	plt.ylabel('y')
+
+	# function to show plot
+	plt.show()
+
+def main(x, y, stat):
+	# observations / data
+
+	# estimating coefficients
+	b = estimate_coef(x, y)
+# 	print("Estimated coefficients:\nb_0 = {} \
+# 		\nb_1 = {}".format(b[0], b[1]))
+
+	# plotting regression line
+	plot_regression_line(x, y, b, stat)
+
+header = lst[0]
+# print(header)
+x = np.array([0])
+y = np.array([0])
+for stat in header[9:]:
+    ind = header.index(stat)
+    x = []
+    y = []
+    for player in lst[1:]:
+#     if type(player[0]) is int:
+        x = np.append(x, float(player[0]))
+#     print(player[14])
+#         print(player[ind])
+        y = np.append(y, float(player[ind]))
+        
+    main(x, y, stat)
+#     print(y)
+    res = stats.linregress(x, y)
+    print(stat + ": " + f"R-squared: {res.rvalue**2:.6f}")
+# main(x, y)
+
+# res = stats.linregress(x, y)
+# print(f"R-squared: {res.rvalue**2:.6f}")

### 2022-11-09 Code/LR_Test_WNBA.py
Uploading code that runs linear regression, WNBA

@@ -1,0 +1,87 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+from scipy import stats
+
+lst = []
+with open("WNBA_Combined.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        lst.append(row)
+
+
+
+# print(x)
+# print(y)
+
+def estimate_coef(x, y):
+	# number of observations/points
+	n = np.size(x)
+
+	# mean of x and y vector
+	m_x = np.mean(x)
+	m_y = np.mean(y)
+
+	# calculating cross-deviation and deviation about x
+	SS_xy = np.sum(y*x) - n*m_y*m_x
+	SS_xx = np.sum(x*x) - n*m_x*m_x
+
+	# calculating regression coefficients
+	b_1 = SS_xy / SS_xx
+	b_0 = m_y - b_1*m_x
+
+	return (b_0, b_1)
+
+def plot_regression_line(x, y, b, stat):
+	# plotting the actual points as scatter plot
+	plt.scatter(x, y, color = "m",
+			marker = "o", s = 30)
+
+	# predicted response vector
+	y_pred = b[0] + b[1]*x
+
+	# plotting the regression line
+	plt.plot(x, y_pred, color = "g")
+
+	# putting labels
+	plt.title(stat)
+	plt.xlabel('x')
+	plt.ylabel('y')
+
+	# function to show plot
+	plt.show()
+
+def main(x, y, stat):
+	# observations / data
+
+	# estimating coefficients
+	b = estimate_coef(x, y)
+# 	print("Estimated coefficients:\nb_0 = {} \
+# 		\nb_1 = {}".format(b[0], b[1]))
+
+	# plotting regression line
+# 	plot_regression_line(x, y, b, stat)
+
+header = lst[0]
+# print(header)
+x = np.array([0])
+y = np.array([0])
+for stat in header[9:]:
+    ind = header.index(stat)
+    x = []
+    y = []
+    for player in lst[1:]:
+#     if type(player[0]) is int:
+        x = np.append(x, float(player[0]))
+#     print(player[14])
+#         print(player[ind])
+        y = np.append(y, float(player[ind]))
+        
+    main(x, y, stat)
+#     print(y)
+    res = stats.linregress(x, y)
+    print(stat + ": " + f"R-squared: {res.rvalue**2:.6f}")
+# main(x, y)
+
+# res = stats.linregress(x, y)
+# print(f"R-squared: {res.rvalue**2:.6f}")

### 2022-11-11 Manuscript/Proposal.tex
Added a lot of research and data

@@
 \usepackage{natbib}
+\usepackage[table]{xcolor}
+\usepackage{booktabs}
 
@@
 In this paper, I hope to compare not just salaries of WNBA vs NBA players, which is something that has been done numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, and statistical performances for each player.
-The goal is to draw conclusions based on the analysis to learn if a similarly performing player in the WNBA gets a comparable salary (adjusted for league revenue/average salaries) to an NBA player, if there are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, and similar questions.
+The goal is to draw conclusions on if there are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, what the correlation is between them, and determine superstar salary proportions (amongst similar questions).
 These are related, yet more in depth questions to the ones posed by Elle Baker.
@@
 The script was also used to combine data for players who played on multiple teams; the original data set would have players listed separately for each team they played on.
-The final data set after filtration and combinated was called NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) for 376 NBA players from the 2017-18 season.
+The final data set after filtration and combination was called NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) for 376 NBA players from the 2017-18 season.
+\par
+Finally, in order to compare the players accurately, the goal of checking their stats and salaries against league revenues must be achieved.
+As such, it was found that the NBA league revenue from 2018 was 8.01 billion USD \cite{NBA_Revenue} and the WNBA league revenue from the same year was 60 million USD \cite{WNBA_Revenue}.
 
 \paragraph{Research Design and Methods}
+Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data.
+We do not need to perform a test to determine if NBA players are paid more than WNBA players, as that is commonly known and very obvious.
+As (INSERT PAPER) discovered, in general WNBA players are actually paid more than NBA players proportional to league revenue.
+In order to confirm that the data for the 2018 seasons fit with this mold, an initial t-test was performed:
+$$H_0 = \mu_0 \leq \mu_1, H_a = \mu_0 > \mu_1$$
+where $\mu_0$ represents the average WNBA salary proportional to league revenue and $\mu_a$ represents the same for the NBA.
+The test was done using a significance level of  5\%, or an alpha value of 0.05, and degrees of freedom $n_1 + n_2 - 2 = 516$.
+By the t-statistic chart, it must be checked if the found t-statistic is greater than 1.646.
+Using a Python script, Comp\_t\_tests.py, it was found that our t-statistic was 16.56, far greater than 1.646, so it can safely be said with 95\% confidence that our data follows the trend found in (INSERT PAPER)
+\par
+The primary focus of the analysis section was to analyze correlation using linear regression between advanced statistics and salary.
+This was done once again using python scripts: LR\_Test\_WNBA and LR\_Test\_NBA, for the WNBA and NBA, respectively, were used to read the data from the cleaned datasets and conduct simple linear regression analysis on all the advanced stats against player salaries.
+Then, using the scipy and matplotlib Python packages, each comparison was graphed with a line of best fit and a correlation coefficient ($R^2$) was found.
+This correlation coefficient tells us how strong the correlation effect is between the statistic and salary, which will help us draw significantly accurate conclusions from the data then if simply the slope was examined.
+There were twelve overlapping statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in both leagues are graphed below:
+
+\begin{tabular}{*5l}    \toprule
+WNBA vs NBA Correlation & Coefficients For & Selected Advanced Statistics \\\midrule
+
+Statistic 
+&  CC: WNBA   & CC: NBA \\
+
+ PER  &  0.019458  & 0.25444 \\
+ FTr   &  0.001641   & 0.033096        \\       
+ ORB\%   &  0.019717   & 0.000274       \\         
+ TRB\%   &  0.000017   & 0.007801        \\
+ AST\%   &  0.041707   & 0.110984         \\
+ STL\%  &  0.003666   & 0.006126        \\    
+ BLK\%  & 0.000444    & 0.000141      \\
+ TOV\%  & 0.010470   & 0.004207     \\
+ USG\%   & 0.033326    & 0.202527        \\
+ OWS   & 0.117693    & 0.337278        \\
+ DWS  & 0.151183    & 0.234252         \\
+ WS   &  0.153999  & 0.364165                  \\
+ (WS/40)/(WS/48)   &  0.047009   & 0.151640        \\
+ \hline
+\end{tabular}
+\newline
+\par
 I plan to conduct numerous hypothesis tests on the questions that I outlined in the introduction 

### 2022-11-13 Code/Comp_t_tests.py
Upload draft of t tests

@@ -1,0 +1,97 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import csv
+from scipy import stats
+import statistics
+
+readNBA = []
+with open("NBA_Combined.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        readNBA.append(row)
+        
+readWNBA = []
+with open("WNBA_Combined.csv", 'r') as file:
+    csvreader = csv.reader(file)
+    for row in csvreader:
+        readWNBA.append(row)
+
+# WNBAheader = ['Player_Salaryy','Player','year_ID','Age','Tm','tm_gms','Tm_Net_Rtg','Pos','G','MP','MP_pct','PER',
+#           'TS_pct','ThrPAr','FTr','ORB_pct','TRB_pct','AST_pct','STL_pct','BLK_pct','TOV_pct','USG_pct','OWS',
+#           'DWS','WS','WS40','Composite_Rating','Wins_Generated']
+
+# NBAheader = ['Salary','Player','Pos','Age','Tm','G','MP','PER','TS%','3PAr','FTr','ORB%','DRB%','TRB%','AST%',
+#           'STL%','BLK%','TOV%','USG%','OWS','DWS','WS','WS/48','OBPM','DBPM','BPM','VORP']
+
+wnba_sal = []
+nba_sal = []
+for player in readWNBA[1:]:
+    wnba_sal.append(float(player[0]))
+for player in readNBA[1:]:
+    nba_sal.append(float(player[0]))
+avg_wnba = sum(wnba_sal)/len(wnba_sal)
+median_wnba = statistics.median(wnba_sal)
+avg_nba = sum(nba_sal)/len(nba_sal)
+median_nba = statistics.median(nba_sal)
+# print(avg_wnba)
+# print(avg_nba)
+base_result = stats.ttest_ind(nba_sal, wnba_sal)
+print("Independent T-Test result for comparing base NBA and WNBA salaries: ")
+print(base_result[0])
+
+prop_wnba_sal = []
+prop_nba_sal = []
+for salary in wnba_sal:
+    sal = salary / 60000000
+    prop_wnba_sal.append(sal)
+for salary in nba_sal:
+    salary = salary / 8010000000
+    prop_nba_sal.append(sal)
+
+# print(prop_wnba_sal[0:10])
+# print(prop_nba_sal[0:10])
+prop_result = stats.ttest_ind(prop_nba_sal, prop_wnba_sal)
[…]

### 2022-11-13 Manuscript/Proposal.tex
Update Proposal.tex

@@
 \begin{document}
-\maketitle
+
+   \begin{center}
+       \vspace*{1cm}
+       \Huge
+       \textbf{How to Earn a Salary}
+
+       \Large
+       \vspace{0.5cm}
+        A Closer Look at WNBA and NBA Player Salaries, the Discrepancies, and What Statistics Might Influence Them
+
+       \vspace{1.5cm}
+
+       \textbf{Colin Sullivan}
+
+       \vfill
+
+       \vfill
+
+       \vspace{0.8cm}
+
+        \Large    
+       Department of Statistics\\
+       University of Connecticut\\
+       Storrs, CT\\
+
+   \end{center}
+\newpage
 
@@
 Then, using the scipy and matplotlib Python packages, each comparison was graphed with a line of best fit and a correlation coefficient ($R^2$) was found.
-This correlation coefficient tells us how strong the correlation effect is between the statistic and salary, which will help us draw significantly accurate conclusions from the data then if simply the slope was examined.
+This correlation coefficient tells us how strong the correlation effect is between the statistic and salary, which will help us draw significantly more accurate conclusions from the data then if simply the slope was examined.
 There were twelve overlapping statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in both leagues are graphed below:
@@
 \par
-I plan to conduct numerous hypothesis tests on the questions that I outlined in the introduction 
-section, using p values and t-tests to compare average statistics and salaries across leagues. 
-These would be independent t tests. I also plan on utilizing bootstrapping to conduct hundreds 
-or thousands of such tests using Python to make sure that I am getting the most accurate results 
-not affected by strange outliers (there are a few). I am considering incorporating some regression 
-in to predict future stats, but I am unsure if that will be feasible with my current data sets.
+The graphs for the strongest correlation in both leagues (both the win shares statistic) are shown below:
+\newline
+\includegraphics[width=.85\textwidth]{NBA_WS_Graph.png}
+\newline
+\includegraphics[width=.85\textwidth]{WNBA_WS_Graph.png}
+
+\par
+Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in each league.
+Instead of looking at simple statistics, work that has been done in (INSERT HOOP DREAMS PAPER), this method used the advanced statistics from the data set.
+Six statistics were selected: Minutes Played (MP), Player Efficiency Rating (PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS).
+The script was written to sort the datasets by these statistics, and the top 10 players in each league in these statistics were extracted.
+Then the salaries were compared to each respective league's median salary, and the average of these top 10 players relations to the average were compiled.
+The entries in the graph can be interpreted as follows: If the value is greater than 1, than the top 10 players in that statistic are paid the value as a factor above league average.
+If the value is less than 1, than the top 10 players in that statistic are paid the value as a factor below league average.
+The graph is below:
+\vspace*{1cm}
+\begin{tabular}{*5l} 
+\toprule
+How Much Compared to Avg & the Top 10 Players Are Paid & For Advanced Stats \\\midrule
+
+Statistic &  WNBA  & NBA \\
+
+ MP  &  1.04008  & 0.40559 \\
+ PER   &  0.92558   & 0.77316        \\       
+ FTr   &  1.05375   & 0.50153       \\         
+ OWS   & 0.94929    & 0.88026        \\
+ DWS  & 0.91109    & 0.76192       \\
+ WS   &  1.02335  & 0.79457                 \\
+ \hline
+\end{tabular}
 

### 2022-11-13 Code/Comp_t_tests.py
Corrected crucial mistake in sorting for statistics

@@ -80,12 +80,15 @@ for stat in WNBA_header[9:]:
         sorted_nba.sort(key=lambda x: float(x[nba_ind]))
 #         print(top10wnba)
 #         print(sorted_nba)
-        for player in sorted_wnba[:10]:
+#         print(sorted_wnba)
+#         print(median_nba)
+        for player in sorted_wnba[-10:]:
 #             print(float(player[0]))
             wnbasum += float(player[0]) / median_wnba
 #             print(wnbasum)
-        for player in sorted_nba[:10]:
+        for player in sorted_nba[-10:]:
             nbasum += float(player[0]) / median_nba
+#             print(float(player[0]) / median_nba)
         wnba_stat_sal_avg = wnbasum / 10
         nba_stat_sal_avg = nbasum / 10
         print("WNBA players in the top 10 in " + str(stat) + " are paid on average " + str(wnba_stat_sal_avg) + " more than league average.")

### 2022-11-14 Manuscript/Manuscript.tex
Update and rename Proposal.tex to Manuscript.tex

--- 2022-11-13 Manuscript/Proposal.tex
@@
 \newpage
+
+\paragraph{Abstract}
+In a world where gender pay gaps have been highlighted in nearly every industry, professional sports are no exception.
+The Women’s National Basketball Association is widely known and accepted to pay very significantly less annually to its players than the male league counterpart, the National Basketball Association.
+There has been a lot of research into this pay gap, and this paper will look to add to the conversation by examining the structure of contracts in relation to performance.
+This will be done using statistics from the 2018 seasons of each respective league, using datasets cleaned up and analyzed using various python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.
 
@@
 Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in each league.
-Instead of looking at simple statistics, work that has been done in (INSERT HOOP DREAMS PAPER), this method used the advanced statistics from the data set.
+Instead of looking at simple statistics, work that has been done in (DiCicio), this method used the advanced statistics from the data set.
 Six statistics were selected: Minutes Played (MP), Player Efficiency Rating (PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS).
@@
 
- MP  &  1.04008  & 0.40559 \\
- PER   &  0.92558   & 0.77316        \\       
- FTr   &  1.05375   & 0.50153       \\         
- OWS   & 0.94929    & 0.88026        \\
- DWS  & 0.91109    & 0.76192       \\
- WS   &  1.02335  & 0.79457                 \\
+ MP  &  1.445  & 5.353 \\
+ PER   &  1.254   & 5.608        \\       
+ FTr   &  1.087   & 2.053       \\         
+ OWS   & 1.445    & 5.666        \\
+ DWS  & 1.435    & 3.462       \\
+ WS   &  1.530  & 5.446                 \\
  \hline
@@
 \paragraph{Discussion}
-I expect to find that the NBA players simply get more money for similar performances as that of WNBA 
-players. I also expect to find that they receive more adjusted compensation, although that remains to 
-be seen and is the basis of the comparison idea. I hope that my work will further the conversation on 
-gender equal pay across sports, an issue as widespread as any and not limited to only basketball. 
-I also hope that my work will provide concrete data results to drive future changes should the leagues 
-attempt to lesser the gap between player compensations. If the investigation goes differently then 
-expected, it will be a very pleasant surprise (although may not bode well for the efforts of female 
-players to increase pay).
+The data found from the initial t-tests supported what had been found in the current literature.
+It was confirmed that in our 2018 data WNBA players are paid significantly less than NBA players, a known fact.
+It was also confirmed with a very high degree of certainty that WNBA players are paid more than NBA players proportional to league revenue (around 60 million for the WNBA, 8 billion for the NBA).
+This confirms the findings of (DiCicio), but was found to be particularly interesting and is in general counterintuitive to the narrative surrounding pay in women’s basketball.
+It also changes the conversation on how the pay gap can be narrowed.
+\par
+The main focus of the research was the regression analysis.
+Simple linear regression was used to create $R^2$ correlation coefficients for every advanced statistic in the WNBA and NBA datasets compared to player salaries.
+Since the base salary and proportional salary in the leagues are so different (as shown in the aforementioned t-tests), it seemed most appropriate to simply see which statistics correlated the strongest to salary increases.
+Table 1 shows the selection of correlation coefficients for the thirteen matching statistics between the two league data sets.
+\par
+Surprisingly, the correlation coefficients for almost every statistic was very low.
+For certain less consequential statistics (like Free Throw rate) this discovery was not shocking, but for more relatively important ones like Win Shares / Minutes it was interesting to see such low correlation between performance and pay.
+That being said, of the most important statistics - Player Efficiency Rating and Win Shares (including the Offensive and Defensive Win Shares sub statistics) - the NBA saw a significant increase in correlation coefficient, with all being above 0.23.
+The WNBA, on the flipside, did not see such an increase: Player Efficiency Rating had an extremely small coefficient, and while WS, OWS and DWS were all higher than other statistics, they still capped out at 0.15, lower than 6 of the NBA statistics coefficients.
+The graphs for the two highest correlations, Win Shares, are shown, and it can be visibly seen that higher performance in the NBA leads to higher pay more often than in the women’s counterpart league.
+\par
+Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic impacts pay.
+As Table 2 shows, there is a much clearer trend in performing in the top 10 in a statistic and a pay increase.
+As the WNBA pay spread is much smaller than the NBA’s it is expected that there will be less of an increase.
+An encouraging sign is found in this table: finishing top 10 in every statistic (with the exception of Free Throw rate, the least consequential of the statistics) led to a significant pay increase in both leagues.
+The NBA consistently saw an increase of over 5 times league median salary, while the WNBA hovered around 1.5 times league median.
+\par
+None of the data found is incredibly shocking, but it adds important information to the current research on the subject.
+For the WNBA to be treated seriously by fans and potential athletes, they must begin to move closer to the NBA formula of greater correlation of performance to player paychecks.
+The WNBA is obviously limited by their revenue, but the product is there; women’s basketball is just as enjoyable to watch as mens, but the players remain treated so extremely differently.
+There is money to be had in women’s basketball, as evidenced by several of the collegiate players at the University of Connecticut making over \$1 million per year in endorsement deals under the new NCAA Name Image Likeness rules.
+To see such insignificant correlation between a statistic like Player Efficiency Rating and pay in the WNBA vs the NBA is eye opening to the gender sport inequality.
 
 \paragraph{Conclusion}
-To summarize: My goal in this paper is to further the discussion on the pay gap between gender 
-in basketball and how this holds back the game as a whole. I hope to do this by manipulating 
-existing data sets to include new stats of my own design, which I will then compare to identical 
-stats in the opposite league using hypothesis tests, t-tests, and bootstrapping (and potentially 
-other tests) to come to conclusions regarding the issue and hopefully be pointed in the right 
-direction to find a solution.
+The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball and primarily to see if the salary models were similar (just dramatically scaled down in the much poorer league).
+It was confirmed that the data for the 2018 season fit with precedent in terms of NBA salaries being much higher and WNBA salaries being higher proportional to league revenue.
+The regression analysis revealed that the correlation coefficients for NBA salaries to selected impactful advanced statistics was in general much higher than that of the WNBA, revealing that the aforementioned salary models were dramatically different.
+It was, however, determined that finishing in the top 10 in said statistics resulted in pay increases over league median, however the NBA’s pay increases were significantly more due to their much larger spread in salaries.
+\par
+Next steps from this data involve researching what goes into league revenue and attempting to determine why it has never been profitable.
+It is important to acknowledge one last time that the WNBA is hamstrung by this lack of revenue, but the contract structures reinforce male dominant stereotypes in an already masculine world with a widespread gender pay gap in nearly every industry.
+Ultimately, the NBA players get rewarded much more for performance than the WNBA, which may motivate them to improve more and bring more excitement and popularity to the league.
+With less monetary motivation to perform, the WNBA continues to sell itself well short of its potential and does not give the players the respect that they deserve.
 

### 2022-11-14 Manuscript/Manuscript.tex
Fixed long lines

@@
 In a world where gender pay gaps have been highlighted in nearly every industry, professional sports are no exception. 
-The Women’s National Basketball Association is widely known and accepted to pay very significantly less annually to its players than the male league counterpart, the National Basketball Association.
-There has been a lot of research into this pay gap, and this paper will look to add to the conversation by examining the structure of contracts in relation to performance.
+The Women’s National Basketball Association is widely known and accepted to pay very significantly less annually to its 
+players than the male league counterpart, the National Basketball Association. There has been a lot of research into this 
+pay gap, and this paper will look to add to the conversation by examining the structure of contracts in relation to performance.
 This will be done using statistics from the 2018 seasons of each respective league, using datasets cleaned up and analyzed using various python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.
@@
 \paragraph{Introduction}
-Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates among fans comparing the female counterpart to the much older, more popular National Basketball Association (NBA).
+Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates 
+among fans comparing the female counterpart to the much older, more popular National Basketball Association (NBA). 
 Women’s basketball has a rich history marred by frequent sexism, fewer opportunities, and significantly less financial support.
 \par
-Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented \cite{Shattering_The_Glass}.
-Due to the gender roles and cultural norms, however, women at the time were believed to be very fragile and unable to play physical sports like basketball at anywhere near the same level as men and thus played under modified rules specially created for them \cite{WNBA_Hist}.
-Those familiar with basketball will know that it is usually played competitively with 5 players on each team, and each player runs the full court to play offense on one end and defend the other team on the other end.
-In 1892, however, women initially played “9 on 9”, with the court divided into three areas: offense, defense, and midcourt.
-Each woman was assigned one of these courts in which they played, such that 3 women on each team were in each section.
-The players could not leave their section of the court to go into the other, essentially creating three separate “3 on 3” games.
-The other major difference from the men’s game was that women could only dribble the ball a limited number of times and could not hold it for more than three seconds, rule changed attributed to the fear that the women players might suffer “nervous fatigue” if the game were too strenuous \cite{WNBA_Hist}.
+Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented 
+\cite{Shattering_The_Glass}. Due to the gender roles and cultural norms, however, women at the time were believed 
+to be very fragile and unable to play physical sports like basketball at anywhere near the same level as men and 
+thus played under modified rules specially created for them \cite{WNBA_Hist}. Those familiar with basketball will 
+know that it is usually played competitively with 5 players on each team, and each player runs the full court to 
+play offense on one end and defend the other team on the other end. In 1892, however, women initially played “9 on 9”, 
+with the court divided into three areas: offense, defense, and midcourt. Each woman was assigned one of these courts 
+in which they played, such that 3 women on each team were in each section. The players could not leave their section 
+of the court to go into the other, essentially creating three separate “3 on 3” games. The other major difference 
+from the men’s game was that women could only dribble the ball a limited number of times and could not hold it for 
+more than three seconds, rule changed attributed to the fear that the women players might suffer “nervous fatigue” 
+if the game were too strenuous \cite{WNBA_Hist}.
 \par
-Everyone now knows that such attitudes towards women’s sports are ridiculous and sexist.
-Certainly there are some physical characteristic differences between men and women in sports, but for every male athlete there is a just as accomplished female one.
-The gradual changing of perspectives on women’s athletics is seen in the evolution of women’s basketball; while men’s basketball has largely stayed the same over the course of its almost 130 year history (key changes including the legalization of the slam dunk and the implementation of the three point line), women’s basketball has shifted slowly to eventually become almost identical.
-The sport was played inter-collegiately for decades beginning in the late 19th century, but did not enter the professional realm until the 1970s.
-The sport became officially recognized at the collegiate level by the NCAA in 1982 (a shockingly late introduction), and the Women’s Basketball League (WBL), founded in 1978 and only lasting three seasons, was the first of several failed women’s professional leagues in the late 20th century.
-In the meantime, the NBA experienced a huge boom in popularity largely due to the emergence of superstars Larry Bird, Magic Johnson, and Michael Jordan.
-The massive increase in NBA reach and revenue directly led the league creating, funding, and supporting the WNBA in 1996.
+Everyone now knows that such attitudes towards women’s sports are ridiculous and sexist. Certainly there are some physical 
+characteristic differences between men and women in sports, but for every male athlete there is a just as accomplished 
+female one. The gradual changing of perspectives on women’s athletics is seen in the evolution of women’s basketball; 
+while men’s basketball has largely stayed the same over the course of its almost 130 year history (key changes including 
+the legalization of the slam dunk and the implementation of the three point line), women’s basketball has shifted slowly 
+to eventually become almost identical. The sport was played inter-collegiately for decades beginning in the late 19th 
+century, but did not enter the professional realm until the 1970s. The sport became officially recognized at the collegiate 
+level by the NCAA in 1982 (a shockingly late introduction), and the Women’s Basketball League (WBL), founded in 1978 and 
+only lasting three seasons, was the first of several failed women’s professional leagues in the late 20th century. In the 
+meantime, the NBA experienced a huge boom in popularity largely due to the emergence of superstars Larry Bird, 
+Magic Johnson, and Michael Jordan. The massive increase in NBA reach and revenue directly led the league creating, 
+funding, and supporting the WNBA in 1996.
 \par
-There are minimal differences between today's women’s basketball and the basketball that men have been playing since inception.
-The WNBA and NCAAW ball is slightly smaller than that used by male players, the three point line is slightly shorter, and the game clock features four 10 minute quarters as opposed to the NBA’s 12 minute quarters, but for all intents and purposes the games are the same.
-The rules are otherwise identical, the athletes are similarly talented, the level of play is incredibly high, and the games are very enjoyable to watch.
-There remains only one massive difference between the two leagues: player salaries.
+There are minimal differences between today's women’s basketball and the basketball that men have been playing since 
+inception. The WNBA and NCAAW ball is slightly smaller than that used by male players, the three point line is slightly 
+shorter, and the game clock features four 10 minute quarters as opposed to the NBA’s 12 minute quarters, but for all 
+intents and purposes the games are the same. The rules are otherwise identical, the athletes are similarly talented, 
+the level of play is incredibly high, and the games are very enjoyable to watch. There remains only one massive 
+difference between the two leagues: player salaries.
 \par
-The difference in player salaries is a huge point of contention in women’s basketball discourse and is often the subject of intense ridicule, and it is easy to see why: top NBA players make upwards of 40 million USD per season, not including endorsement deals and other income, while the WNBA most valuable player makes less than the average office worker in New York (an exaggeration, maybe, but Elena Delle Donne made around 120 thousand USD in her MVP year with little extra endorsement money or exposure).
-An NBA rookie signs one contract and has enough wealth to live comfortably off for the rest of his life, often for generations; a WNBA prospect is encouraged to continue her academic studies to ensure that she is able to get a good job after her career is over.
-Now, granted, the league revenue is much much less than that of the NBA, but many questions still remain that this paper hopes to address: Are the players paid fairly compared to league revenue?
-Are they paid fairly compared to NBA players based on basketball production?
+The difference in player salaries is a huge point of contention in women’s basketball discourse and is often the subject 
+of intense ridicule, and it is easy to see why: top NBA players make upwards of 40 million USD per season, not including 
+endorsement deals and other income, while the WNBA most valuable player makes less than the average office worker in 
+New York (an exaggeration, maybe, but Elena Delle Donne made around 120 thousand USD in her MVP year with little extra 
+endorsement money or exposure). An NBA rookie signs one contract and has enough wealth to live comfortably off for 
+the rest of his life, often for generations; a WNBA prospect is encouraged to continue her academic studies to 
+ensure that she is able to get a good job after her career is over. Now, granted, the league revenue is much 
+much less than that of the NBA, but many questions still remain that this paper hopes to address: Are the players 
+paid fairly compared to league revenue? Are they paid fairly compared to NBA players based on basketball production? 
 How can this change?
@@
 \par
-In this paper, I hope to compare not just salaries of WNBA vs NBA players, which is something that has been done numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, and statistical performances for each player.
-The goal is to draw conclusions on if there are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, what the correlation is between them, and determine superstar salary proportions (amongst similar questions).
-These are related, yet more in depth questions to the ones posed by Elle Baker.
+In this paper, I hope to compare not just salaries of WNBA vs NBA players, which is something that has been done 
+numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, 
+and statistical performances for each player. The goal is to draw conclusions on if there are certain statistics in 
+the NBA that result in higher adjusted salaries than the WNBA and vice versa, what the correlation is between them, 
+and determine superstar salary proportions (amongst similar questions). These are related, yet more in depth questions 
+to the ones posed by Elle Baker.
 \par
-The paper will begin by discussing the datasets being analyzed and how they were cleaned and prepped for analysis using Python scripts.
-In the next section, statistical analysis will be performed to try and answer some of the questions posed in this introduction.
-Following that will be a discussion on the results and hypotheses for best paths forward.
+The paper will begin by discussing the datasets being analyzed and how they were cleaned and prepped for analysis using 
+Python scripts. In the next section, statistical analysis will be performed to try and answer some of the questions 
+posed in this introduction. Following that will be a discussion on the results and hypotheses for best paths forward. 
 Finally, there will be a conclusions section to wrap up the paper and review the results.
@@
 \paragraph{Data}
-There are two primary data sets that I will be using to conduct the analysis in this paper.
-The first of these is a database of historical WNBA data dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed on his GitHub account \cite{first}.
-Unlike most data sets, this one contains a multitude of advanced statistics for each player.
-The statistical categories are as follows:
+There are two primary data sets that I will be using to conduct the analysis in this paper. The first of these is a 
+database of historical WNBA data dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed 
+on his GitHub account \cite{first}. Unlike most data sets, this one contains a multitude of advanced statistics for 
+each player. The statistical categories are as follows:
 \newline
 \par
-Player Name, Year, Age, Team, Team Games, Team Net Rating, Position, Games, Minutes Played, Minutes Played (as a percentage of team total minutes), Player Efficiency Rating, True Shooting \%, Free Throw Rating, Offensive Rebounding \%, Total Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, Win Shares per 40 games, Composite Rating, and Wins Generated.
+Player Name, Year, Age, Team, Team Games, Team Net Rating, Position, Games, Minutes Played, Minutes Played (as a percentage 
+of team total minutes), Player Efficiency Rating, True Shooting \%, Free Throw Rating, Offensive Rebounding \%, Total 
+Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, 
+Win Shares, Win Shares per 40 games, Composite Rating, and Wins Generated.
 \newline
 \par
-Using another data set from Paine’s Github, a Python script called WNBA\_Cleanup.py was written and used to combine salary data from the two sources and filter out the desired candidate players from the data set.
-This paper will focus on comparing players from the 2018 NBA and WNBA seasons, so that filter was applied.
-The script also filtered out players who played less than 5 games and those that did not have salary data, so as to make the set as standardized as possible and remove outliers.
-The final data set after filtration and combination was called WNBA\_Combined.csv and has data (the aforementioned statistical categories) for 166 WNBA players from the 2018 season.
+Using another data set from Paine’s Github, a Python script called WNBA\_Cleanup.py was written and used to combine 
+salary data from the two sources and filter out the desired candidate players from the data set. This paper will 
+focus on comparing players from the 2018 NBA and WNBA seasons, so that filter was applied. The script also 
+filtered out players who played less than 5 games and those that did not have salary data, so as to make the set as 
+standardized as possible and remove outliers. The final data set after filtration and combination was called 
+WNBA\_Combined.csv and has data (the aforementioned statistical categories) for 166 WNBA players from the 2018 season.
 \par
-A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site Kaggle \cite{nba}.
-This data set contains a list of all players from the NBA in the 2017-18 season, with similar advanced statistics to the WNBA dataset:
+A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site 
+Kaggle \cite{nba}. This data set contains a list of all players from the NBA in the 2017-18 season, with similar 
+advanced statistics to the WNBA dataset:
 \newline
 \par
-Player Name, Position, Age, Team, Games Played, Minutes Played, Player Efficiency Rating, True Shooting \%, 3 Point Attempt Rate, Free Throw Rate, Offensive Rebounding \%, Defensive Rebounding \%, Total Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, Win Shares per 48 games, Offensive Box +/-, Defensive Box +/-, Box +/-, and VORP.
+Player Name, Position, Age, Team, Games Played, Minutes Played, Player Efficiency Rating, True Shooting \%, 
+3 Point Attempt Rate, Free Throw Rate, Offensive Rebounding \%, Defensive Rebounding \%, Total Rebounding \%, 
+Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, 
+Win Shares per 48 games, Offensive Box +/-, Defensive Box +/-, Box +/-, and VORP.
 \newline
 \par
-The data set notably did not contain player salaries, so a second NBA data set was necessary.
-One was found on GitHub, created by Erik Gregory Webb, which contains the player salaries for every NBA player from 2000-2020 \cite{nba_salaries}.
-In a similar manner to the WNBA data, a Python script called NBA\_Cleanup.py was created to combine the data from the two datasets into one csv file.
-The data was then cleaned up by removing outliers (players playing under 5 games) and those that did not have salary data.
-The script was also used to combine data for players who played on multiple teams; the original data set would have players listed separately for each team they played on.
-The final data set after filtration and combination was called NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) for 376 NBA players from the 2017-18 season.
+The data set notably did not contain player salaries, so a second NBA data set was necessary. One was found on GitHub, 
+created by Erik Gregory Webb, which contains the player salaries for every NBA player from 2000-2020 \cite{nba_salaries}. 
+In a similar manner to the WNBA data, a Python script called NBA\_Cleanup.py was created to combine the data from the two 
+datasets into one csv file. The data was then cleaned up by removing outliers (players playing under 5 games) and those 
+that did not have salary data. The script was also used to combine data for players who played on multiple teams; 
+the original data set would have players listed separately for each team they played on. The final data set after filtration 
+and combination was called NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) 
+for 376 NBA players from the 2017-18 season.
 \par
-Finally, in order to compare the players accurately, the goal of checking their stats and salaries against league revenues must be achieved.
-As such, it was found that the NBA league revenue from 2018 was 8.01 billion USD \cite{NBA_Revenue} and the WNBA league revenue from the same year was 60 million USD \cite{WNBA_Revenue}.
+Finally, in order to compare the players accurately, the goal of checking their stats and salaries against league 
+revenues must be achieved. As such, it was found that the NBA league revenue from 2018 was 8.01 billion USD \cite{NBA_Revenue} 
+and the WNBA league revenue from the same year was 60 million USD \cite{WNBA_Revenue}.
 
 \paragraph{Research Design and Methods}
-Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data.
-We do not need to perform a test to determine if NBA players are paid more than WNBA players, as that is commonly known and very obvious.
-As (INSERT PAPER) discovered, in general WNBA players are actually paid more than NBA players proportional to league revenue.
-In order to confirm that the data for the 2018 seasons fit with this mold, an initial t-test was performed:
+Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data. We do 
+not need to perform a test to determine if NBA players are paid more than WNBA players, as that is commonly known and 
+very obvious. As (INSERT PAPER) discovered, in general WNBA players are actually paid more than NBA players proportional 
+to league revenue. In order to confirm that the data for the 2018 seasons fit with this mold, an initial t-test was performed:
 $$H_0 = \mu_0 \leq \mu_1, H_a = \mu_0 > \mu_1$$
-where $\mu_0$ represents the average WNBA salary proportional to league revenue and $\mu_a$ represents the same for the NBA.
-The test was done using a significance level of  5\%, or an alpha value of 0.05, and degrees of freedom $n_1 + n_2 - 2 = 516$.
-By the t-statistic chart, it must be checked if the found t-statistic is greater than 1.646.
-Using a Python script, Comp\_t\_tests.py, it was found that our t-statistic was 16.56, far greater than 1.646, so it can safely be said with 95\% confidence that our data follows the trend found in (INSERT PAPER)
+where $\mu_0$ represents the average WNBA salary proportional to league revenue and $\mu_a$ represents the same for 
+the NBA. The test was done using a significance level of  5\%, or an alpha value of 0.05, and degrees of freedom 
+$n_1 + n_2 - 2 = 516$. By the t-statistic chart, it must be checked if the found t-statistic is greater than 1.646. 
+Using a Python script, Comp\_t\_tests.py, it was found that our t-statistic was 16.56, far greater than 1.646, 
+so it can safely be said with 95\% confidence that our data follows the trend found in (INSERT PAPER)
 \par
-The primary focus of the analysis section was to analyze correlation using linear regression between advanced statistics and salary.
-This was done once again using python scripts: LR\_Test\_WNBA and LR\_Test\_NBA, for the WNBA and NBA, respectively, were used to read the data from the cleaned datasets and conduct simple linear regression analysis on all the advanced stats against player salaries.
-Then, using the scipy and matplotlib Python packages, each comparison was graphed with a line of best fit and a correlation coefficient ($R^2$) was found.
-This correlation coefficient tells us how strong the correlation effect is between the statistic and salary, which will help us draw significantly more accurate conclusions from the data then if simply the slope was examined.
-There were twelve overlapping statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in both leagues are graphed below:
+The primary focus of the analysis section was to analyze correlation using linear regression between advanced statistics 
+and salary. This was done once again using python scripts: LR\_Test\_WNBA and LR\_Test\_NBA, for the WNBA and NBA, 
+respectively, were used to read the data from the cleaned datasets and conduct simple linear regression analysis on 
+all the advanced stats against player salaries. Then, using the scipy and matplotlib Python packages, each comparison 
+was graphed with a line of best fit and a correlation coefficient ($R^2$) was found. This correlation coefficient 
+tells us how strong the correlation effect is between the statistic and salary, which will help us draw significantly 
+more accurate conclusions from the data then if simply the slope was examined. There were twelve overlapping 
+statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in both leagues
+ are graphed below:
 
@@
 \par
-Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in each league.
-Instead of looking at simple statistics, work that has been done in (DiCicio), this method used the advanced statistics from the data set.
-Six statistics were selected: Minutes Played (MP), Player Efficiency Rating (PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS).
-The script was written to sort the datasets by these statistics, and the top 10 players in each league in these statistics were extracted.
-Then the salaries were compared to each respective league's median salary, and the average of these top 10 players relations to the average were compiled.
-The entries in the graph can be interpreted as follows: If the value is greater than 1, than the top 10 players in that statistic are paid the value as a factor above league average.
-If the value is less than 1, than the top 10 players in that statistic are paid the value as a factor below league average.
-The graph is below:
+Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in 
+each league. Instead of looking at simple statistics, work that has been done in (DiCicio), this method used the 
+advanced statistics from the data set. Six statistics were selected: Minutes Played (MP), Player Efficiency Rating 
+(PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS). 
+The script was written to sort the datasets by these statistics, and the top 10 players in each league in these 
+statistics were extracted. Then the salaries were compared to each respective league's median salary, and the 
+average of these top 10 players relations to the average were compiled. The entries in the graph can be interpreted 
+as follows: If the value is greater than 1, than the top 10 players in that statistic are paid the value as a factor 
+above league average. If the value is less than 1, than the top 10 players in that statistic are paid the value as a 
+factor below league average. The graph is below:
 \vspace*{1cm}
@@
 \paragraph{Discussion}
-The data found from the initial t-tests supported what had been found in the current literature.
-It was confirmed that in our 2018 data WNBA players are paid significantly less than NBA players, a known fact.
-It was also confirmed with a very high degree of certainty that WNBA players are paid more than NBA players proportional to league revenue (around 60 million for the WNBA, 8 billion for the NBA).
-This confirms the findings of (DiCicio), but was found to be particularly interesting and is in general counterintuitive to the narrative surrounding pay in women’s basketball.
-It also changes the conversation on how the pay gap can be narrowed.
+The data found from the initial t-tests supported what had been found in the current literature. It was confirmed that in 
+our 2018 data WNBA players are paid significantly less than NBA players, a known fact. It was also confirmed with a very 
+high degree of certainty that WNBA players are paid more than NBA players proportional to league revenue (around 60 
+million for the WNBA, 8 billion for the NBA). This confirms the findings of (DiCicio), but was found to be particularly 
+interesting and is in general counterintuitive to the narrative surrounding pay in women’s basketball. It also changes 
+the conversation on how the pay gap can be narrowed.
 \par
-The main focus of the research was the regression analysis.
-Simple linear regression was used to create $R^2$ correlation coefficients for every advanced statistic in the WNBA and NBA datasets compared to player salaries.
-Since the base salary and proportional salary in the leagues are so different (as shown in the aforementioned t-tests), it seemed most appropriate to simply see which statistics correlated the strongest to salary increases.
-Table 1 shows the selection of correlation coefficients for the thirteen matching statistics between the two league data sets.
+The main focus of the research was the regression analysis. Simple linear regression was used to create $R^2$ 
+correlation coefficients for every advanced statistic in the WNBA and NBA datasets compared to player salaries. 
+Since the base salary and proportional salary in the leagues are so different (as shown in the aforementioned t-tests), 
+it seemed most appropriate to simply see which statistics correlated the strongest to salary increases. Table 1 shows 
+the selection of correlation coefficients for the thirteen matching statistics between the two league data sets.
 \par
-Surprisingly, the correlation coefficients for almost every statistic was very low.
-For certain less consequential statistics (like Free Throw rate) this discovery was not shocking, but for more relatively important ones like Win Shares / Minutes it was interesting to see such low correlation between performance and pay.
-That being said, of the most important statistics - Player Efficiency Rating and Win Shares (including the Offensive and Defensive Win Shares sub statistics) - the NBA saw a significant increase in correlation coefficient, with all being above 0.23.
-The WNBA, on the flipside, did not see such an increase: Player Efficiency Rating had an extremely small coefficient, and while WS, OWS and DWS were all higher than other statistics, they still capped out at 0.15, lower than 6 of the NBA statistics coefficients.
-The graphs for the two highest correlations, Win Shares, are shown, and it can be visibly seen that higher performance in the NBA leads to higher pay more often than in the women’s counterpart league.
+Surprisingly, the correlation coefficients for almost every statistic was very low. For certain less consequential 
+statistics (like Free Throw rate) this discovery was not shocking, but for more relatively important ones like Win 
+Shares / Minutes it was interesting to see such low correlation between performance and pay. That being said, of the 
+most important statistics - Player Efficiency Rating and Win Shares (including the Offensive and Defensive Win Shares 
+sub statistics) - the NBA saw a significant increase in correlation coefficient, with all being above 0.23. The WNBA, 
+on the flipside, did not see such an increase: Player Efficiency Rating had an extremely small coefficient, and while 
+WS, OWS and DWS were all higher than other statistics, they still capped out at 0.15, lower than 6 of the NBA statistics 
+coefficients. The graphs for the two highest correlations, Win Shares, are shown, and it can be visibly seen that higher
+ performance in the NBA leads to higher pay more often than in the women’s counterpart league.
 \par
-Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic impacts pay.
-As Table 2 shows, there is a much clearer trend in performing in the top 10 in a statistic and a pay increase.
-As the WNBA pay spread is much smaller than the NBA’s it is expected that there will be less of an increase.
-An encouraging sign is found in this table: finishing top 10 in every statistic (with the exception of Free Throw rate, the least consequential of the statistics) led to a significant pay increase in both leagues.
-The NBA consistently saw an increase of over 5 times league median salary, while the WNBA hovered around 1.5 times league median.
+Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic 
+impacts pay. As Table 2 shows, there is a much clearer trend in performing in the top 10 in a statistic and a pay increase. 
+As the WNBA pay spread is much smaller than the NBA’s it is expected that there will be less of an increase. An encouraging 
+sign is found in this table: finishing top 10 in every statistic (with the exception of Free Throw rate, the least 
+consequential of the statistics) led to a significant pay increase in both leagues. The NBA consistently saw an increase 
+of over 5 times league median salary, while the WNBA hovered around 1.5 times league median.
 \par
 None of the data found is incredibly shocking, but it adds important information to the current research on the subject. 
-For the WNBA to be treated seriously by fans and potential athletes, they must begin to move closer to the NBA formula of greater correlation of performance to player paychecks.
-The WNBA is obviously limited by their revenue, but the product is there; women’s basketball is just as enjoyable to watch as mens, but the players remain treated so extremely differently.
-There is money to be had in women’s basketball, as evidenced by several of the collegiate players at the University of Connecticut making over \$1 million per year in endorsement deals under the new NCAA Name Image Likeness rules.
-To see such insignificant correlation between a statistic like Player Efficiency Rating and pay in the WNBA vs the NBA is eye opening to the gender sport inequality.
+For the WNBA to be treated seriously by fans and potential athletes, they must begin to move closer to the NBA formula 
+of greater correlation of performance to player paychecks. The WNBA is obviously limited by their revenue, but the product 
+is there; women’s basketball is just as enjoyable to watch as mens, but the players remain treated so extremely differently. 
+There is money to be had in women’s basketball, as evidenced by several of the collegiate players at the University of 
+Connecticut making over \$1 million per year in endorsement deals under the new NCAA Name Image Likeness rules. To see 
+such insignificant correlation between a statistic like Player Efficiency Rating and pay in the WNBA vs the NBA is eye
+ opening to the gender sport inequality.
 
 \paragraph{Conclusion}
-The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball and primarily to see if the salary models were similar (just dramatically scaled down in the much poorer league).
-It was confirmed that the data for the 2018 season fit with precedent in terms of NBA salaries being much higher and WNBA salaries being higher proportional to league revenue.
-The regression analysis revealed that the correlation coefficients for NBA salaries to selected impactful advanced statistics was in general much higher than that of the WNBA, revealing that the aforementioned salary models were dramatically different.
-It was, however, determined that finishing in the top 10 in said statistics resulted in pay increases over league median, however the NBA’s pay increases were significantly more due to their much larger spread in salaries.
+The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball 
+and primarily to see if the salary models were similar (just dramatically scaled down in the much poorer league). 
+It was confirmed that the data for the 2018 season fit with precedent in terms of NBA salaries being much higher 
+and WNBA salaries being higher proportional to league revenue. The regression analysis revealed that the 
+correlation coefficients for NBA salaries to selected impactful advanced statistics was in general much higher 
+than that of the WNBA, revealing that the aforementioned salary models were dramatically different. It was, 
+however, determined that finishing in the top 10 in said statistics resulted in pay increases over league median, 
+however the NBA’s pay increases were significantly more due to their much larger spread in salaries. 
 \par
-Next steps from this data involve researching what goes into league revenue and attempting to determine why it has never been profitable.
-It is important to acknowledge one last time that the WNBA is hamstrung by this lack of revenue, but the contract structures reinforce male dominant stereotypes in an already masculine world with a widespread gender pay gap in nearly every industry.
-Ultimately, the NBA players get rewarded much more for performance than the WNBA, which may motivate them to improve more and bring more excitement and popularity to the league.
-With less monetary motivation to perform, the WNBA continues to sell itself well short of its potential and does not give the players the respect that they deserve.
+Next steps from this data involve researching what goes into league revenue and attempting to determine why it 
+has never been profitable. It is important to acknowledge one last time that the WNBA is hamstrung by this lack 
+of revenue, but the contract structures reinforce male dominant stereotypes in an already masculine world with a 
+widespread gender pay gap in nearly every industry. Ultimately, the NBA players get rewarded much more for 
+performance than the WNBA, which may motivate them to improve more and bring more excitement and popularity to the league. 
+With less monetary motivation to perform, the WNBA continues to sell itself well short of its potential and does 
+not give the players the respect that they deserve.
 

### 2022-11-14 Code/figure.R
Delete figure.R

@@ -1,4 +1,0 @@
-pdf("Manuscript/cars.pdf", pointsize = 9, height = 4, width = 6)
-plot(dist ~ speed, data = cars,
-     xlab = "speed (mph)", ylab = "distance (feet)")
-dev.off()

### 2022-11-14 Code/table.R
Delete table.R

@@ -1,9 +1,0 @@
-library(xtable)
-
-n <- 10
-dat <- data.frame(normal = rnorm(n),
-                  poisson = rpois(n, lambda = 4),
-                  gamma = rgamma(n, scale = 2, shape = 2))
-
-xtab <- xtable(dat, digits = c(0, 3, 0, 3))
-print(xtab, include.rownames = FALSE)

### 2022-11-23 Manuscript/Manuscript.tex
Updated table with dataset variables

--- 2022-11-14 Manuscript/Manuscript.tex
@@
 % highlighting hyper links
-\usepackage[colorlinks=true, citecolor=blue]{hyperref}
+\usepackage[colorlinks=true, citepcolor=blue]{hyperref}
 
@@
 
-\paragraph{Abstract}
+\begin{abstract}
 In a world where gender pay gaps have been highlighted in nearly every industry, professional sports are no exception.
@@
 This will be done using statistics from the 2018 seasons of each respective league, using datasets cleaned up and analyzed using various python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.
+\end{abstract}
 
-\paragraph{Introduction}
+\section{Introduction}
 Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates among fans comparing the female counterpart to the much older, more popular National Basketball Association (NBA).
@@
 \par
-Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented \cite{Shattering_The_Glass}.
-Due to the gender roles and cultural norms, however, women at the time were believed to be very fragile and unable to play physical sports like basketball at anywhere near the same level as men and thus played under modified rules specially created for them \cite{WNBA_Hist}.
+Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented \citep{Shattering_The_Glass}.
+Due to the gender roles and cultural norms, however, women at the time were believed to be very fragile and unable to play physical sports like basketball at anywhere near the same level as men and thus played under modified rules specially created for them \citep{WNBA_Hist}.
 Those familiar with basketball will know that it is usually played competitively with 5 players on each team, and each player runs the full court to play offense on one end and defend the other team on the other end.
@@
 The players could not leave their section of the court to go into the other, essentially creating three separate “3 on 3” games.
-The other major difference from the men’s game was that women could only dribble the ball a limited number of times and could not hold it for more than three seconds, rule changed attributed to the fear that the women players might suffer “nervous fatigue” if the game were too strenuous \cite{WNBA_Hist}.
+The other major difference from the men’s game was that women could only dribble the ball a limited number of times and could not hold it for more than three seconds, rule changed attributed to the fear that the women players might suffer “nervous fatigue” if the game were too strenuous \citep{WNBA_Hist}.
 \par
@@
 how WNBA players, despite getting paid far less, have much less "salary inequality"
-\cite{baker2020comparison}. Another argues for an increase in salaries in the WNBA (a cause everyone
-should be able to get behind) by citing the league's growth in fans and revenues \cite{ettienne2019s}.
+\citep{}{baker2020comparison}. Another argues for an increase in salaries in the WNBA (a cause everyone
+should be able to get behind) by citing the league's growth in fans and revenues \citep{ettienne2019s}.
 There seems to be a lack of specific adjusted salary comparisons, however, and I hope to build off articles
@@
 
-\paragraph{Data}
+\section{Data}
 There are two primary data sets that I will be using to conduct the analysis in this paper.
-The first of these is a database of historical WNBA data dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed on his GitHub account \cite{first}.
+The first of these is a database of historical WNBA data dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed on his GitHub account \citep{first}.
 Unlike most data sets, this one contains a multitude of advanced statistics for each player.
-The statistical categories are as follows:
-\newline
-\par
-Player Name, Year, Age, Team, Team Games, Team Net Rating, Position, Games, Minutes Played, Minutes Played (as a percentage of team total minutes), Player Efficiency Rating, True Shooting \%, Free Throw Rating, Offensive Rebounding \%, Total Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, Win Shares per 40 games, Composite Rating, and Wins Generated.
-\newline
+
 \par
@@
 \par
-A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site Kaggle \cite{nba}.
-This data set contains a list of all players from the NBA in the 2017-18 season, with similar advanced statistics to the WNBA dataset:
+A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site Kaggle \citep{nba}.
+This data set contains a list of all players from the NBA in the 2017-18 season, with similar advanced statistics to the WNBA dataset.
+The statistics found in the datasets can be seen in the following table:
 \newline
 \par
-Player Name, Position, Age, Team, Games Played, Minutes Played, Player Efficiency Rating, True Shooting \%, 3 Point Attempt Rate, Free Throw Rate, Offensive Rebounding \%, Defensive Rebounding \%, Total Rebounding \%, Assist \%, Steal \%, Block \%, Turnover \%, Usage \%, Offensive Win Shares, Defensive Win Shares, Win Shares, Win Shares per 48 games, Offensive Box +/-, Defensive Box +/-, Box +/-, and VORP.
+\begin{tabular}{*5l} 
+\toprule
+Statistics From the Datasets \\\midrule
+
+Common Statistics & WNBA Exclusive & NBA Exclusive\\ 
+\hline
+ Player Name  &  Team Net Rating & 3 Point Attempt Rate  \\
+ Age &  Composite Rating & Offensive Box +/-      \\ 
+ Position &  Wins Generated & Defensive Box +/-\\
+ Team &  & Box +/-     \\         
+ Minutes Played &  & VORP       \\
+ Player Efficiency Rating & &      \\
+ True Shooting \% & & \\
+ Free Throw Rate & &\\
+ Offensive Rebounding \% & & \\
+ Total Rebounding \% & & \\
+ Assist \% & & \\
+ Steal \% & & \\
+ Block \% & & \\
+ Turnover \% & & \\
+ Usage \% & & \\
+ Offensive Win Shares & & \\
+ Defensive Win Shares  & & \\
+ Win Shares & & \\
+ Win Shares per 40 games & & \\
+ Salary & & \\
+
+ \hline
+\end{tabular}
 \newline
@@
 The data set notably did not contain player salaries, so a second NBA data set was necessary.
-One was found on GitHub, created by Erik Gregory Webb, which contains the player salaries for every NBA player from 2000-2020 \cite{nba_salaries}.
+One was found on GitHub, created by Erik Gregory Webb, which contains the player salaries for every NBA player from 2000-2020 \citep{nba_salaries}.
 In a similar manner to the WNBA data, a Python script called NBA\_Cleanup.py was created to combine the data from the two datasets into one csv file.
@@
 Finally, in order to compare the players accurately, the goal of checking their stats and salaries against league revenues must be achieved.
-As such, it was found that the NBA league revenue from 2018 was 8.01 billion USD \cite{NBA_Revenue} and the WNBA league revenue from the same year was 60 million USD \cite{WNBA_Revenue}.
+As such, it was found that the NBA league revenue from 2018 was 8.01 billion USD \citep{NBA_Revenue} and the WNBA league revenue from the same year was 60 million USD \citep{WNBA_Revenue}.
 
-\paragraph{Research Design and Methods}
+\section{Research Design and Methods}
 Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data.
@@
 
-\paragraph{Discussion}
+\section{Discussion}
 The data found from the initial t-tests supported what had been found in the current literature.
@@
 
-\paragraph{Conclusion}
+\section{Conclusion}
 The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball and primarily to see if the salary models were similar (just dramatically scaled down in the much poorer league).
@@
 It is important to acknowledge one last time that the WNBA is hamstrung by this lack of revenue, but the contract structures reinforce male dominant stereotypes in an already masculine world with a widespread gender pay gap in nearly every industry.
-Ultimately, the NBA players get rewarded much more for performance than the WNBA, which may motivate them to improve more and bring more excitement and popularity to the league.
+Ultimately, the NBA players get rewarded much more for performance than the WNBA, which may motivate them to improve more and bring more excitepment and popularity to the league.
 With less monetary motivation to perform, the WNBA continues to sell itself well short of its potential and does not give the players the respect that they deserve.
+
+\section{Supplementary Material}
+If interested, the data sets and Python scripts used can be found in the GitHub repo \url{https://github.com/statds/course-paper-or-presentation-ColinSullivan9}.
 

### 2022-12-01 Manuscript/Manuscript.tex
Update Manuscript.tex

@@
 % highlighting hyper links
-\usepackage[colorlinks=true, citepcolor=blue]{hyperref}
+\usepackage[colorlinks=true, citecolor=blue]{hyperref}
 
@@
 \begin{abstract}
-In a world where gender pay gaps have been highlighted in nearly every industry, professional sports are no exception.
-The Women’s National Basketball Association is widely known and accepted to pay very significantly less annually to its players than the male league counterpart, the National Basketball Association.
-There has been a lot of research into this pay gap, and this paper will look to add to the conversation by examining the structure of contracts in relation to performance.
-This will be done using statistics from the 2018 seasons of each respective league, using datasets cleaned up and analyzed using various python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.
+In a world where gender pay gaps have been highlighted in nearly every industry, professional sports are no exception. The Women’s National Basketball 
+Association is widely known and accepted to pay very significantly less annually to its players than the male league counterpart, the National Basketball 
+Association. There has been a lot of research into this pay gap, and this paper will look to add to the conversation by examining the structure of 
+contracts in relation to performance. This will be done using statistics from the 2018 seasons of each respective league, using datasets cleaned up and 
+analyzed with various python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.	
 \end{abstract}
@@
 \section{Introduction}
-Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates among fans comparing the female counterpart to the much older, more popular National Basketball Association (NBA).
-Women’s basketball has a rich history marred by frequent sexism, fewer opportunities, and significantly less financial support.
+Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates among fans comparing the female 
+league to the much older, more popular National Basketball Association (NBA). Women’s basketball has a rich history marred by frequent sexism, fewer 
+opportunities, and significantly less financial support.
 \par
-Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented \citep{Shattering_The_Glass}.
-Due to the gender roles and cultural norms, however, women at the time were believed to be very fragile and unable to play physical sports like basketball at anywhere near the same level as men and thus played under modified rules specially created for them \citep{WNBA_Hist}.
-Those familiar with basketball will know that it is usually played competitively with 5 players on each team, and each player runs the full court to play offense on one end and defend the other team on the other end.
-In 1892, however, women initially played “9 on 9”, with the court divided into three areas: offense, defense, and midcourt.
-Each woman was assigned one of these courts in which they played, such that 3 women on each team were in each section.
-The players could not leave their section of the court to go into the other, essentially creating three separate “3 on 3” games.
-The other major difference from the men’s game was that women could only dribble the ball a limited number of times and could not hold it for more than three seconds, rule changed attributed to the fear that the women players might suffer “nervous fatigue” if the game were too strenuous \citep{WNBA_Hist}.
+Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented \citep{Shattering_The_Glass}. Due to the 
+gender roles and cultural norms of the time, however, women at the time were believed to be very fragile and unable to play physical sports like 
+basketball at anywhere near the same level as men. As such, women played under modified rules specially created for them \citep{WNBA_Hist}. Those familiar 
+with basketball will know that it is usually played competitively with 5 players on each team, and each player runs the full court to play offense on one 
+end and defend the other team on the other end. In 1892, however, women initially played “9 on 9”, with the court divided into three areas: offense, 
+defense, and mid-court. Each player was assigned one of these courts in which they played, such that 3 women on each team were in each section. The 
+players could not leave their section of the court to go into the other, essentially creating three separate “3 on 3” games. The other major rule change 
+from Naismith's game was that women could only dribble the ball a limited number of times and could not hold it for more than three seconds, a rule 
+attributed to the fear that the women players might suffer “nervous fatigue” if the game were too strenuous \citep{WNBA_Hist}.
 \par
-Everyone now knows that such attitudes towards women’s sports are ridiculous and sexist.
-Certainly there are some physical characteristic differences between men and women in sports, but for every male athlete there is a just as accomplished female one.
-The gradual changing of perspectives on women’s athletics is seen in the evolution of women’s basketball; while men’s basketball has largely stayed the same over the course of its almost 130 year history (key changes including the legalization of the slam dunk and the implementation of the three point line), women’s basketball has shifted slowly to eventually become almost identical.
-The sport was played inter-collegiately for decades beginning in the late 19th century, but did not enter the professional realm until the 1970s.
-The sport became officially recognized at the collegiate level by the NCAA in 1982 (a shockingly late introduction), and the Women’s Basketball League (WBL), founded in 1978 and only lasting three seasons, was the first of several failed women’s professional leagues in the late 20th century.
-In the meantime, the NBA experienced a huge boom in popularity largely due to the emergence of superstars Larry Bird, Magic Johnson, and Michael Jordan.
-The massive increase in NBA reach and revenue directly led the league creating, funding, and supporting the WNBA in 1996.
+The gradual changing of perspectives on women’s athletics is seen in the evolution of women’s basketball; while the game of men’s basketball has largely 
+stayed the same over the course of its almost 130 year history (key changes including the legalization of the slam dunk and the implementation of the 
+three point line), women’s basketball has shifted slowly to eventually become almost identical. The sport was played inter-collegiately for decades 
+beginning in the late 19th century, but did not enter the professional realm until the 1970s. The sport became officially recognized at the collegiate 
+level by the NCAA in 1982 (a shockingly late introduction), and the Women’s Basketball League (WBL), founded in 1978 and only lasting three seasons, was 
+the first of several failed women’s professional leagues in the late 20th century. Meanwhile, the NBA experienced a huge boom in popularity largely due to 
+the emergence of superstars Larry Bird, Magic Johnson, and Michael Jordan. The massive increase in NBA reach and revenue directly led to the league 
+creating, funding, and supporting the WNBA in 1996.
 \par
-There are minimal differences between today's women’s basketball and the basketball that men have been playing since inception.
-The WNBA and NCAAW ball is slightly smaller than that used by male players, the three point line is slightly shorter, and the game clock features four 10 minute quarters as opposed to the NBA’s 12 minute quarters, but for all intents and purposes the games are the same.
-The rules are otherwise identical, the athletes are similarly talented, the level of play is incredibly high, and the games are very enjoyable to watch.
-There remains only one massive difference between the two leagues: player salaries.
+There are minimal differences between today's women’s basketball and the basketball that men have been playing since inception. The WNBA and wome's 
+collegiate ball is slightly smaller than that used by male players, the three point line is a foot to two feet shorter, and the game clock features 
+four 10 minute quarters as opposed to the NBA’s 12 minute quarters. Nevertheless, for all intents and purposes the games are the same. The rules are otherwise 
+identical, the athletes are similarly talented, the level of play is incredibly high, and the games are very enjoyable to watch. There remains only one 
+massive difference between the two leagues: player salaries.
 \par
-The difference in player salaries is a huge point of contention in women’s basketball discourse and is often the subject of intense ridicule, and it is easy to see why: top NBA players make upwards of 40 million USD per season, not including endorsement deals and other income, while the WNBA most valuable player makes less than the average office worker in New York (an exaggeration, maybe, but Elena Delle Donne made around 120 thousand USD in her MVP year with little extra endorsement money or exposure).
-An NBA rookie signs one contract and has enough wealth to live comfortably off for the rest of his life, often for generations; a WNBA prospect is encouraged to continue her academic studies to ensure that she is able to get a good job after her career is over.
-Now, granted, the league revenue is much much less than that of the NBA, but many questions still remain that this paper hopes to address: Are the players paid fairly compared to league revenue?
-Are they paid fairly compared to NBA players based on basketball production?
-How can this change?
+The difference in player salaries is a huge point of contention in women’s basketball discourse and is often the subject of intense ridicule, and it is 
+easy to see why: top NBA players make upwards of 40 million USD per season, not including endorsement deals and other income, while the WNBA most valuable 
+player makes less than the average office worker in New York (an exaggeration, maybe, but Elena Delle Donne made around 120 thousand USD in her MVP year 
+with little extra endorsement money or exposure). An NBA rookie signs one contract and has enough wealth to live comfortably off for the rest of his life, 
+often for generations; a WNBA prospect is encouraged to continue her academic studies to ensure that she is able to get a well-paying job after her career 
+is over. Now, granted, the league revenue is much much less than that of the NBA, but many questions still remain that this paper hopes to address: Are 
+the players paid fairly compared to league revenue? Are they compensated equally based on basketball production? How can this change?
 \par
@@
 how WNBA players, despite getting paid far less, have much less "salary inequality"
-\citep{}{baker2020comparison}. Another argues for an increase in salaries in the WNBA (a cause everyone
-should be able to get behind) by citing the league's growth in fans and revenues \citep{ettienne2019s}.
-There seems to be a lack of specific adjusted salary comparisons, however, and I hope to build off articles
-like these to further the conversation.
+\citep{baker2020comparison}. Another argues for an increase in salaries in the WNBA (a cause everyone
+should be able to get behind) by citing the league's growth in fans and revenues \citep{ettienne2019s}. A third paper, by Ball State statistics student 
+Alyssa Bridge took a close look at media tie ins to the WNBA and found that, among other inequalities, only 2\% of ESPN’s SportsCenter airtime is devoted 
+to discussing it. \citep{Bridge}
+This paper hopes to build off some of the research done by Hailey DiCicco in "Hoop Dreams: An Empirical Analysis of the Gender Wage Gap in Professional 
+Basketball". DiCicio performed regression analysis on the correlation between win percentage and team performance in basic statistics (rebounds or steals, 
+for example). She was, however, unable to come to any significant conclusions beyond that of basic comparisons of salaries \citep{Hoop_Dreams}.
 \par
-In this paper, I hope to compare not just salaries of WNBA vs NBA players, which is something that has been done numerous times, but to go further in depth and examine salaries vs team salary caps, league revenues, games played, and statistical performances for each player.
-The goal is to draw conclusions on if there are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, what the correlation is between them, and determine superstar salary proportions (amongst similar questions).
-These are related, yet more in depth questions to the ones posed by Elle Baker.
+This paper will not compare just salaries of WNBA vs NBA players, which is something that has been done numerous times, but will go further in depth and 
+examine the correlation between individual player performance and performance in numerous advanced statistics. The goal is to draw conclusions on if there 
+are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, what the correlation is between them, and 
+determine "superstar" salary factors.
 \par
-The paper will begin by discussing the datasets being analyzed and how they were cleaned and prepped for analysis using Python scripts.
-In the next section, statistical analysis will be performed to try and answer some of the questions posed in this introduction.
-Following that will be a discussion on the results and hypotheses for best paths forward.
-Finally, there will be a conclusions section to wrap up the paper and review the results.
+Immediately following this section, this paper will discuss the datasets being analyzed and how they were cleaned and prepped for analysis using Python 
+scripts. In the next section, statistical analysis will be performed to try and answer some of the questions posed in this introduction. Following that 
+will be a discussion on the results and hypotheses for best paths forward. Finally, there will be a conclusions section to wrap up the paper and review 
+the results.
 
 \section{Data}
-There are two primary data sets that I will be using to conduct the analysis in this paper.
-The first of these is a database of historical WNBA data dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed on his GitHub account \citep{first}.
-Unlike most data sets, this one contains a multitude of advanced statistics for each player.
+There are two primary data sets that I will be using to conduct the analysis in this paper. The first of these is a database of historical WNBA data 
+dating from 2001-2020, collected by Neil Payne, sports editor at 538, and displayed on his GitHub account \citep{first}. Unlike most data sets, this one 
+contains a multitude of advanced statistics for each player.
 
 \par
-Using another data set from Paine’s Github, a Python script called WNBA\_Cleanup.py was written and used to combine salary data from the two sources and filter out the desired candidate players from the data set.
-This paper will focus on comparing players from the 2018 NBA and WNBA seasons, so that filter was applied.
-The script also filtered out players who played less than 5 games and those that did not have salary data, so as to make the set as standardized as possible and remove outliers.
-The final data set after filtration and combination was called WNBA\_Combined.csv and has data (the aforementioned statistical categories) for 166 WNBA players from the 2018 season.
+Using another data set from Paine’s Github, a Python script called WNBA\_Cleanup.py was written and used to combine salary data from the two sources and 
+filter out the desired candidate players from the data set. This paper will focus on comparing players from the 2018 NBA and WNBA seasons, so that filter 
+was applied. The script also filtered out players who played less than 5 games and those that did not have salary data, so as to make the set as 
+standardized as possible and remove outliers. The final data set after filtration and combination was called WNBA\_Combined.csv and has data (the 
+aforementioned statistical categories) for 166 WNBA players from the 2018 season.
 \par
-A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site Kaggle \citep{nba}.
-This data set contains a list of all players from the NBA in the 2017-18 season, with similar advanced statistics to the WNBA dataset.
-The statistics found in the datasets can be seen in the following table:
+A corresponding NBA dataset was found, this time belonging to Mustafa Baris Camli on the data-sharing site Kaggle \citep{nba}. This data set contains a 
+list of all players from the NBA in the 2017-18 season, with similar advanced statistics to the WNBA dataset. The statistics found in the datasets can be 
+seen in the following table:
 \newline
@@
 \par
-The data set notably did not contain player salaries, so a second NBA data set was necessary.
-One was found on GitHub, created by Erik Gregory Webb, which contains the player salaries for every NBA player from 2000-2020 \citep{nba_salaries}.
-In a similar manner to the WNBA data, a Python script called NBA\_Cleanup.py was created to combine the data from the two datasets into one csv file.
-The data was then cleaned up by removing outliers (players playing under 5 games) and those that did not have salary data.
-The script was also used to combine data for players who played on multiple teams; the original data set would have players listed separately for each team they played on.
-The final data set after filtration and combination was called NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) for 376 NBA players from the 2017-18 season.
+The data set notably did not contain player salaries, so a second NBA data set was necessary. One was found on GitHub, created by Erik Gregory Webb, which 
+contains the player salaries for every NBA player from 2000-2020 \citep{nba_salaries}. In a similar manner to the WNBA data, a Python script called 
+NBA\_Cleanup.py was created to combine the data from the two datasets into one csv file. The data was then cleaned up by removing outliers (players 
+playing under 5 games) and those that did not have salary data. The script was also used to combine data for players who played on multiple teams; the 
+original data set would have players listed separately for each team they played on. The final data set after filtration and combination was called 
+NBA\_Combined.csv and has data (the aforementioned statistical categories, plus player salaries) for 376 NBA players from the 2017-18 season.
 \par
-Finally, in order to compare the players accurately, the goal of checking their stats and salaries against league revenues must be achieved.
-As such, it was found that the NBA league revenue from 2018 was 8.01 billion USD \citep{NBA_Revenue} and the WNBA league revenue from the same year was 60 million USD \citep{WNBA_Revenue}.
+Finally, in order to compare the players accurately, the goal of checking their stats and salaries against league revenues must be achieved. As such, it 
+was found that the NBA league revenue from 2018 was 8.01 billion USD \citep{NBA_Revenue} and the WNBA league revenue from the same year was 60 million USD 
+\citep{WNBA_Revenue}.
 
 \section{Research Design and Methods}
-Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data.
-We do not need to perform a test to determine if NBA players are paid more than WNBA players, as that is commonly known and very obvious.
-As (INSERT PAPER) discovered, in general WNBA players are actually paid more than NBA players proportional to league revenue.
-In order to confirm that the data for the 2018 seasons fit with this mold, an initial t-test was performed:
-$$H_0 = \mu_0 \leq \mu_1, H_a = \mu_0 > \mu_1$$
-where $\mu_0$ represents the average WNBA salary proportional to league revenue and $\mu_a$ represents the same for the NBA.
-The test was done using a significance level of  5\%, or an alpha value of 0.05, and degrees of freedom $n_1 + n_2 - 2 = 516$.
-By the t-statistic chart, it must be checked if the found t-statistic is greater than 1.646.
-Using a Python script, Comp\_t\_tests.py, it was found that our t-statistic was 16.56, far greater than 1.646, so it can safely be said with 95\% confidence that our data follows the trend found in (INSERT PAPER)
+Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data. We do not need to perform a test to 
+determine if NBA players are paid more than WNBA players, as that is commonly known and very obvious. As $\textit{Forbes}$ basketball author David Berri 
+discovered, in general WNBA players are still paid significantly less than NBA players proportional to league revenue \citep{WNBA_Revenue}. In order to 
+confirm that the data for the 2018 seasons fit with this mold, an initial t-test was performed:
+$$H_0 = \mu_1 \leq \mu_0, H_a = \mu_1 > \mu_0$$
+where $\mu_0$ represents the average WNBA salary proportional to league revenue and $\mu_a$ represents the same for the NBA. The test was done using a 
+significance level of  5\%, or an alpha value of 0.05, and degrees of freedom $n_1 + n_2 - 2 = 516$. By the t-statistic chart, it must be checked if the 
+found t-statistic is greater than 1.646. Using a Python script, Comp\_t\_tests.py, it was found that our t-statistic was 16.56. Since we want to convert 
+this into a one tailed test, we halve the t-statistic to get 8.28, which is far greater than 1.646, so it can safely be said with 95\% confidence that our 
+data follows the trend found in Berri's article.
 \par
-The primary focus of the analysis section was to analyze correlation using linear regression between advanced statistics and salary.
-This was done once again using python scripts: LR\_Test\_WNBA and LR\_Test\_NBA, for the WNBA and NBA, respectively, were used to read the data from the cleaned datasets and conduct simple linear regression analysis on all the advanced stats against player salaries.
-Then, using the scipy and matplotlib Python packages, each comparison was graphed with a line of best fit and a correlation coefficient ($R^2$) was found.
-This correlation coefficient tells us how strong the correlation effect is between the statistic and salary, which will help us draw significantly more accurate conclusions from the data then if simply the slope was examined.
-There were twelve overlapping statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in both leagues are graphed below:
+The primary focus of the analysis section was to analyze correlation using linear regression between advanced statistics and salary. This was done once 
+again using python scripts: LR\_Test\_WNBA and LR\_Test\_NBA, for the WNBA and NBA, respectively, were used to read the data from the cleaned datasets and 
+conduct simple linear regression analysis on all the advanced stats against player salaries. Then, using the scipy and matplotlib Python packages, each 
+comparison was graphed with a line of best fit and a correlation coefficient ($R^2$) was found. This correlation coefficient tells us how strong the 
+correlation effect is between the statistic and salary, which will help us draw significantly more accurate conclusions from the data then if simply the 
+slope was examined. There were twelve overlapping statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in 
+both leagues are graphed below, with the standard error in parenthesis:
 
@@
 
- PER  &  0.019458  & 0.25444 \\
- FTr   &  0.001641   & 0.033096        \\       
- ORB\%   &  0.019717   & 0.000274       \\         
- TRB\%   &  0.000017   & 0.007801        \\
- AST\%   &  0.041707   & 0.110984         \\
- STL\%  &  0.003666   & 0.006126        \\    
- BLK\%  & 0.000444    & 0.000141      \\
- TOV\%  & 0.010470   & 0.004207     \\
- USG\%   & 0.033326    & 0.202527        \\
- OWS   & 0.117693    & 0.337278        \\
- DWS  & 0.151183    & 0.234252         \\
- WS   &  0.153999  & 0.364165                  \\
- (WS/40)/(WS/48)   &  0.047009   & 0.151640        \\
+ PER  &  0.019458 (1.576)  & 0.25444 (0.014) \\
+ FTr   &  0.001641 (0.036)  & 0.033096 (0.009)       \\ 
+ ORB\%   &  0.019717 (0.929)  & 0.000274 (0.273)       \\   
+ TRB\%   &  0.000017 (1.080)   & 0.007801 (0.337)       \\
+ AST\%   &  0.041707 (1.936)  & 0.110984 (0.608)    \\
+ STL\%  &  0.003666 (0.175)  & 0.006126 (0.045)       \\    
+ BLK\%  & 0.000444 (0.337)   & 0.000141 (0.114)     \\
+ TOV\%  & 0.010470 (1.432)  & 0.004207 (0.315)    \\
+ USG\%   & 0.033326 (1.146)   & 0.202527 (0.350)       \\
+ OWS   & 0.117693 (0.276)   & 0.337278 (0.121)        \\
+ DWS  & 0.151183 (0.133)   & 0.234252 (0.068)         \\
+ WS   &  0.153999 (0.365)  & 0.364165 (0.164)                 \\
+ (WS/40)/(WS/48)   &  0.047009 (0.022)  & 0.151640 (0.004) \\
  \hline
@@
 \par
-Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in each league.
-Instead of looking at simple statistics, work that has been done in (DiCicio), this method used the advanced statistics from the data set.
-Six statistics were selected: Minutes Played (MP), Player Efficiency Rating (PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS).
-The script was written to sort the datasets by these statistics, and the top 10 players in each league in these statistics were extracted.
-Then the salaries were compared to each respective league's median salary, and the average of these top 10 players relations to the average were compiled.
-The entries in the graph can be interpreted as follows: If the value is greater than 1, than the top 10 players in that statistic are paid the value as a factor above league average.
-If the value is less than 1, than the top 10 players in that statistic are paid the value as a factor below league average.
-The graph is below:
+Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in each league. Instead of looking at 
+simple statistics, work that has been done in (DiCicio), this method used the advanced statistics from the data set. Six statistics were selected: Minutes 
+Played (MP), Player Efficiency Rating (PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS). The 
+script was written to sort the datasets by these statistics, and the top 10 players in each league in these statistics were extracted. Then the salaries 
+were compared to each respective league's median salary, and the average of these top 10 players relations to the average were compiled. The entries in 
+the graph can be interpreted as follows: If the value is greater than 1, than the top 10 players in that statistic are paid the value as a factor above 
+league average. If the value is less than 1, than the top 10 players in that statistic are paid the value as a factor below league average. The graph is 
+below:
 \vspace*{1cm}
@@
 \section{Discussion}
-The data found from the initial t-tests supported what had been found in the current literature.
-It was confirmed that in our 2018 data WNBA players are paid significantly less than NBA players, a known fact.
-It was also confirmed with a very high degree of certainty that WNBA players are paid more than NBA players proportional to league revenue (around 60 million for the WNBA, 8 billion for the NBA).
-This confirms the findings of (DiCicio), but was found to be particularly interesting and is in general counterintuitive to the narrative surrounding pay in women’s basketball.
-It also changes the conversation on how the pay gap can be narrowed.
+The data found from the initial t-tests supported what had been found in the current literature. It was confirmed that in our 2018 data WNBA players are 
+paid significantly less than NBA players, a known fact. It was also confirmed with a very high degree of certainty that WNBA players are paid more than 
+NBA players proportional to league revenue (around 60 million for the WNBA, 8 billion for the NBA). This confirms the findings of (DiCicio), but was found 
+to be particularly interesting and is in general counterintuitive to the narrative surrounding pay in women’s basketball. It also changes the conversation 
+on how the pay gap can be narrowed.
 \par
-The main focus of the research was the regression analysis.
-Simple linear regression was used to create $R^2$ correlation coefficients for every advanced statistic in the WNBA and NBA datasets compared to player salaries.
-Since the base salary and proportional salary in the leagues are so different (as shown in the aforementioned t-tests), it seemed most appropriate to simply see which statistics correlated the strongest to salary increases.
-Table 1 shows the selection of correlation coefficients for the thirteen matching statistics between the two league data sets.
+The main focus of the research was the regression analysis. Simple linear regression was used to create $R^2$ correlation coefficients for every advanced 
+statistic in the WNBA and NBA datasets compared to player salaries. Since the base salary and proportional salary in the leagues are so different (as 
+shown in the aforementioned t-tests), it seemed most appropriate to simply see which statistics correlated the strongest to salary increases. Table 1 
+shows the selection of correlation coefficients for the thirteen matching statistics between the two league data sets.
 \par
-Surprisingly, the correlation coefficients for almost every statistic was very low.
-For certain less consequential statistics (like Free Throw rate) this discovery was not shocking, but for more relatively important ones like Win Shares / Minutes it was interesting to see such low correlation between performance and pay.
-That being said, of the most important statistics - Player Efficiency Rating and Win Shares (including the Offensive and Defensive Win Shares sub statistics) - the NBA saw a significant increase in correlation coefficient, with all being above 0.23.
-The WNBA, on the flipside, did not see such an increase: Player Efficiency Rating had an extremely small coefficient, and while WS, OWS and DWS were all higher than other statistics, they still capped out at 0.15, lower than 6 of the NBA statistics coefficients.
-The graphs for the two highest correlations, Win Shares, are shown, and it can be visibly seen that higher performance in the NBA leads to higher pay more often than in the women’s counterpart league.
+Surprisingly, the correlation coefficients for almost every statistic was very low. For certain less consequential statistics (like Free Throw rate) this 
+discovery was not shocking, but for more relatively important ones like Win Shares / Minutes it was interesting to see such low correlation between 
+performance and pay. That being said, of the most important statistics - Player Efficiency Rating and Win Shares (including the Offensive and Defensive 
+Win Shares sub statistics) - the NBA saw a significant increase in correlation coefficient, with all being above 0.23. The WNBA, on the flipside, did not 
+see such an increase: Player Efficiency Rating had an extremely small coefficient, and while WS, OWS and DWS were all higher than other statistics, they 
+still capped out at 0.15, lower than 6 of the NBA statistics coefficients. The graphs for the two highest correlations, Win Shares, are shown, and it can 
+be visibly seen that higher performance in the NBA leads to higher pay more often than in the women’s counterpart league.
 \par
-Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic impacts pay.
-As Table 2 shows, there is a much clearer trend in performing in the top 10 in a statistic and a pay increase.
-As the WNBA pay spread is much smaller than the NBA’s it is expected that there will be less of an increase.
-An encouraging sign is found in this table: finishing top 10 in every statistic (with the exception of Free Throw rate, the least consequential of the statistics) led to a significant pay increase in both leagues.
-The NBA consistently saw an increase of over 5 times league median salary, while the WNBA hovered around 1.5 times league median.
+The visual distribution of pay vs performance in the graphs is particularly interesting. While the correlation coefficient for both the WNBA and the NBA 
+when compared to Win Shared is relatively high, the NBA graph looks much more traditionally distributed: Examining it finds the data generally following 
+the trend line. The WNBA graph, however, looks much less traditional. It forms a sort of "U" shape, where players are almost all clustered into two salary 
+buckets which are distributed similarly over the y axis. While there are more "higher performing" players in the higher pay bucket, the distribution is 
+nevertheless strange to see.
 \par
-None of the data found is incredibly shocking, but it adds important information to the current research on the subject.
-For the WNBA to be treated seriously by fans and potential athletes, they must begin to move closer to the NBA formula of greater correlation of performance to player paychecks.
-The WNBA is obviously limited by their revenue, but the product is there; women’s basketball is just as enjoyable to watch as mens, but the players remain treated so extremely differently.
-There is money to be had in women’s basketball, as evidenced by several of the collegiate players at the University of Connecticut making over \$1 million per year in endorsement deals under the new NCAA Name Image Likeness rules.
-To see such insignificant correlation between a statistic like Player Efficiency Rating and pay in the WNBA vs the NBA is eye opening to the gender sport inequality.
+Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic impacts pay. As Table 2 shows, 
+there is a much clearer trend in performing in the top 10 in a statistic and a pay increase. As the WNBA pay spread is much smaller than the NBA’s it is 
+expected that there will be less of an increase. An encouraging sign is found in this table: finishing top 10 in every statistic (with the exception of 
+Free Throw rate, the least consequential of the statistics) led to a significant pay increase in both leagues. The NBA consistently saw an increase of 
+over 5 times league median salary, while the WNBA hovered around 1.5 times league median.
+\par
+None of the data found is incredibly shocking, but it adds important information to the current research on the subject. For the WNBA to be treated 
+seriously by fans and potential athletes, they must begin to move closer to the NBA formula of greater correlation of performance to player paychecks. The 
+WNBA is obviously limited by their revenue, but the product is there; women’s basketball is just as enjoyable to watch as mens, but the players remain 
+treated so extremely differently. There is money to be had in women’s basketball, as evidenced by several of the collegiate players at the University of 
+Connecticut making over \$1 million per year in endorsement deals under the new NCAA Name Image Likeness rules. To see such insignificant correlation 
+between a statistic like Player Efficiency Rating and pay in the WNBA vs the NBA is eye opening to the gender sport inequality.
 
 \section{Conclusion}
-The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball and primarily to see if the salary models were similar (just dramatically scaled down in the much poorer league).
-It was confirmed that the data for the 2018 season fit with precedent in terms of NBA salaries being much higher and WNBA salaries being higher proportional to league revenue.
-The regression analysis revealed that the correlation coefficients for NBA salaries to selected impactful advanced statistics was in general much higher than that of the WNBA, revealing that the aforementioned salary models were dramatically different.
-It was, however, determined that finishing in the top 10 in said statistics resulted in pay increases over league median, however the NBA’s pay increases were significantly more due to their much larger spread in salaries.
+The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball and primarily to see if the salary 
+models were similar (just dramatically scaled down in the much poorer league). It was confirmed that the data for the 2018 season fit with precedent 
+in terms of NBA salaries being much higher, both base and proportional to league revenue. The regression analysis revealed that the correlation 
+coefficients for NBA salaries to selected impactful advanced statistics was in general much higher than that of the WNBA, revealing that the 
+aforementioned salary models were dramatically different. It was, however, determined that finishing in the top 10 in said statistics resulted in pay 
+increases over league median, however the NBA’s pay increases were significantly more due to their much larger spread in salaries. 
+\par
+Perhaps one of the most interesting conclusions is drawn when examining the graphs of NBA and WNBA pay distribution against Win Shares performance. 
+Not only is there much less pay increase opportunity in the WNBA, but it seems as though almost all players are grouped into one of two pay rates, 
+which may or may not be based solely upon performance.
 \par
 Next steps from this data involve researching what goes into league revenue and attempting to determine why it has never been profitable. 
-It is important to acknowledge one last time that the WNBA is hamstrung by this lack of revenue, but the contract structures reinforce male dominant stereotypes in an already masculine world with a widespread gender pay gap in nearly every industry.
-Ultimately, the NBA players get rewarded much more for performance than the WNBA, which may motivate them to improve more and bring more excitepment and popularity to the league.
-With less monetary motivation to perform, the WNBA continues to sell itself well short of its potential and does not give the players the respect that they deserve.
+It is important to acknowledge one last time that the WNBA is hamstrung by this lack of revenue, but the contract structures and much lower 
+pay proportional to league revenue reinforce male dominant stereotypes in an already masculine world with a widespread gender pay gap in nearly
+every industry. Ultimately, the NBA players get rewarded much more for performance than the WNBA, which may motivate them to improve more and bring 
+more excitepment and popularity to the league. With less monetary motivation to perform, the WNBA continues to sell itself well short of its potential 
+and does not give the players the respect that they deserve.
 

### 2022-12-05 Manuscript/Manuscript.tex
Finalized Manuscript

@@
 contracts in relation to performance. This will be done using statistics from the 2018 seasons of each respective league, using datasets cleaned up and 
-analyzed with various python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.	
+analyzed with various Python scripts to eventually draw conclusions on the relation between performance and pay for male and female athletes.	
 \end{abstract}
@@
 \section{Introduction}
+\hspace*{10mm}
 Since the formulation of the Women’s National Basketball Association (WNBA) in 1996, there have been never-ending debates among fans comparing the female 
@@
 Women’s basketball began being played in 1892 at Smith College, just one year after men’s basketball was invented \citep{Shattering_The_Glass}. Due to the 
-gender roles and cultural norms of the time, however, women at the time were believed to be very fragile and unable to play physical sports like 
+gender roles and cultural norms of the time, however, women were believed to be very fragile and unable to play physical sports like 
 basketball at anywhere near the same level as men. As such, women played under modified rules specially created for them \citep{WNBA_Hist}. Those familiar 
 with basketball will know that it is usually played competitively with 5 players on each team, and each player runs the full court to play offense on one 
-end and defend the other team on the other end. In 1892, however, women initially played “9 on 9”, with the court divided into three areas: offense, 
-defense, and mid-court. Each player was assigned one of these courts in which they played, such that 3 women on each team were in each section. The 
+end and defend the other team on the other. In 1892, however, women initially played “9 on 9”, with the court divided into three areas: offense, 
+defense, and mid-court. Each player was assigned to play within one of these cours, such that 3 women on each team were in each section. The 
 players could not leave their section of the court to go into the other, essentially creating three separate “3 on 3” games. The other major rule change 
 from Naismith's game was that women could only dribble the ball a limited number of times and could not hold it for more than three seconds, a rule 
-attributed to the fear that the women players might suffer “nervous fatigue” if the game were too strenuous \citep{WNBA_Hist}.
+attributed to the fear that the players might suffer “nervous fatigue” if the game were too strenuous \citep{WNBA_Hist}.
 \par
@@
 three point line), women’s basketball has shifted slowly to eventually become almost identical. The sport was played inter-collegiately for decades 
-beginning in the late 19th century, but did not enter the professional realm until the 1970s. The sport became officially recognized at the collegiate 
+beginning in the late 19th century, but did not enter the professional realm until the 1970s. Women's basketball became officially recognized at the collegiate 
 level by the NCAA in 1982 (a shockingly late introduction), and the Women’s Basketball League (WBL), founded in 1978 and only lasting three seasons, was 
@@
 \par
-There are minimal differences between today's women’s basketball and the basketball that men have been playing since inception. The WNBA and wome's 
+There are minimal differences between today's women’s basketball and the basketball that men have been playing since inception. The WNBA and women's 
 collegiate ball is slightly smaller than that used by male players, the three point line is a foot to two feet shorter, and the game clock features 
@@
 This paper will not compare just salaries of WNBA vs NBA players, which is something that has been done numerous times, but will go further in depth and 
-examine the correlation between individual player performance and performance in numerous advanced statistics. The goal is to draw conclusions on if there 
+examine the correlation between individual player pay and performance in numerous advanced statistics. The goal is to draw conclusions on if there 
 are certain statistics in the NBA that result in higher adjusted salaries than the WNBA and vice versa, what the correlation is between them, and 
@@
 \section{Data}
+\hspace*{10mm}
 There are two primary data sets that I will be using to conduct the analysis in this paper. The first of these is a database of historical WNBA data 
@@
 \section{Research Design and Methods}
-Initially, we began by performing a series of paired sample t-tests to confirm some expectations from the data. We do not need to perform a test to 
+\hspace*{10mm}
+Initially, we began by performing a paired sample t-test to confirm some expectations from the data. We do not need to perform a test to 
 determine if NBA players are paid more than WNBA players, as that is commonly known and very obvious. As $\textit{Forbes}$ basketball author David Berri 
@@
 slope was examined. There were twelve overlapping statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in 
-both leagues are graphed below, with the standard error in parenthesis:
+both leagues are shown below, with the standard error in parenthesis:
 
@@
  \hline
+ Sample Size & 142 & 376
 \end{tabular}
@@
 Finally, within the Comp\_t\_tests.py Python script, analysis was conducted on the salaries of the top players in each league. Instead of looking at 
-simple statistics, work that has been done in (DiCicio), this method used the advanced statistics from the data set. Six statistics were selected: Minutes 
+simple statistics, work that has been done by DiCicio \citep{Hoop_Dreams}, this method used the advanced statistics from the data set. Six statistics were selected: Minutes 
 Played (MP), Player Efficiency Rating (PER), Free Throw rate (FTr), Offensive Win Shares (OWS), Defensive Win Shares (DWS), and Win Shares (WS). The 
@@
 the graph can be interpreted as follows: If the value is greater than 1, than the top 10 players in that statistic are paid the value as a factor above 
-league average. If the value is less than 1, than the top 10 players in that statistic are paid the value as a factor below league average. The graph is 
+league average. If the value is less than 1, than the top 10 players in that statistic are paid the value as a factor below league average. The chart is 
 below:
@@
 \section{Discussion}
-The data found from the initial t-tests supported what had been found in the current literature. It was confirmed that in our 2018 data WNBA players are 
-paid significantly less than NBA players, a known fact. It was also confirmed with a very high degree of certainty that WNBA players are paid more than 
-NBA players proportional to league revenue (around 60 million for the WNBA, 8 billion for the NBA). This confirms the findings of (DiCicio), but was found 
-to be particularly interesting and is in general counterintuitive to the narrative surrounding pay in women’s basketball. It also changes the conversation 
-on how the pay gap can be narrowed.
+\hspace*{10mm}
+The data found from the initial t-test supported what had been found in the current literature. It was confirmed that in our 2018 data WNBA players are 
+paid significantly less than NBA players, a known fact. It was also confirmed with a very high degree of certainty that WNBA players are paid less than 
+NBA players proportional to league revenue (around 60 million for the WNBA, 8 billion for the NBA). This confirms the findings of David Berri \citep{WNBA_Revenue}, and was found 
+to be particularly interesting while being generally intuitive based upon the narrative surrounding pay in women’s basketball.
 \par
@@
 The visual distribution of pay vs performance in the graphs is particularly interesting. While the correlation coefficient for both the WNBA and the NBA 
-when compared to Win Shared is relatively high, the NBA graph looks much more traditionally distributed: Examining it finds the data generally following 
+when compared to Win Shares is relatively high, the NBA graph looks much more traditionally distributed: Examining it finds the data generally following 
 the trend line. The WNBA graph, however, looks much less traditional. It forms a sort of "U" shape, where players are almost all clustered into two salary 
@@
 \section{Conclusion}
+\hspace*{10mm}
 The motivation of this paper was to try and determine what goes into the pay gap in men’s and women’s basketball and primarily to see if the salary 

### 2022-12-05 Manuscript/Manuscript.tex
Update Manuscript.tex

@@
-\documentclass[12pt]{article}
+\documentclass[12pt, titlepage]{article}
 
@@
 \begin{document}
+\maketitle
 
-   \begin{center}
-       \vspace*{1cm}
-       \Huge
-       \textbf{How to Earn a Salary}
-
-       \Large
-       \vspace{0.5cm}
-        A Closer Look at WNBA and NBA Player Salaries, the Discrepancies, and What Statistics Might Influence Them
-
-       \vspace{1.5cm}
-
-       \textbf{Colin Sullivan}
-
-       \vfill
-
-       \vfill
-
-       \vspace{0.8cm}
-
-        \Large    
-       Department of Statistics\\
-       University of Connecticut\\
-       Storrs, CT\\
-
-   \end{center}
+%   \begin{center}
+%       \vspace*{1cm}
+%       \Huge
+%       \textbf{How to Earn a Salary}
+%
+%       \Large
+%       \vspace{0.5cm}
+%        A Closer Look at WNBA and NBA Player Salaries, the Discrepancies, and What Statistics Might Influence Them
+%            
+%       \vspace{1.5cm}
+%
+%       \textbf{Colin Sullivan}
+%       
+%       \vfill
+%
+%
+%       \vfill
+%            
+%            
+%       \vspace{0.8cm}
+%     
+%         
+%        \Large    
+%       Department of Statistics\\
+%       University of Connecticut\\
+%       Storrs, CT\\
+%            
+%   \end{center}
 \newpage
@@
 list of all players from the NBA in the 2017-18 season, with similar advanced statistics to the WNBA dataset. The statistics found in the datasets can be 
-seen in the following table:
-\newline
-\par
-\begin{tabular}{*5l} 
+seen in \ref{tab:stats}.
+
+\begin{table}[tbp]
+    \caption{Statistics From the Datasets}
+    \label{tab:stats}
+\centering
+\begin{tabular}{crr}
 \toprule
-Statistics From the Datasets \\\midrule
 
 Common Statistics & WNBA Exclusive & NBA Exclusive\\ 
-\hline
+\midrule
  Player Name  &  Team Net Rating & 3 Point Attempt Rate  \\
@@
 
- \hline
+ \bottomrule
 \end{tabular}
-\newline
+\end{table}
+
 \par
@@
 slope was examined. There were twelve overlapping statistics between the WNBA and NBA datasets, and the correlation coefficients for these statistics in 
-both leagues are shown below, with the standard error in parenthesis:
+both leagues are shown in \ref{tab:main_data}, with the standard error in parenthesis.
 
-\begin{tabular}{*5l}    \toprule
-WNBA vs NBA Correlation & Coefficients For & Selected Advanced Statistics \\\midrule
+\begin{table}[tbp]
+    \caption{WNBA vs NBA Correlation Coefficients For Selected Advanced Statistics}
+    \label{tab:main_data}
+\centering
+\begin{tabular}{crr}
+\toprule
 
-Statistic 
-&  CC: WNBA   & CC: NBA \\
-
+Statistic &  CC: WNBA   & CC: NBA \\ 
+\midrule
  PER  &  0.019458 (1.576)  & 0.25444 (0.014) \\
@@
  \hline
- Sample Size & 142 & 376
+ Sample Size & 142 & 376\\
+ \bottomrule
 \end{tabular}
-\newline
+\end{table}
+
 \par
-The graphs for the strongest correlation in both leagues (both the win shares statistic) are shown below:
-\newline
-\includegraphics[width=.85\textwidth]{NBA_WS_Graph.png}
-\newline
-\includegraphics[width=.85\textwidth]{WNBA_WS_Graph.png}
+The graphs for the strongest correlation in both leagues (both the win shares statistic) are shown in figures \ref{fig:NBA_WS} and \ref{fig:WNBA_WS}.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=.75\textwidth]{NBA_WS_Graph.png}
+  \caption{NBA Player Salaries vs Win Share Performance.}
+  \label{fig:NBA_WS}
+\end{figure}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=.75\textwidth]{WNBA_WS_Graph.png}
+  \caption{WNBA Player Salaries vs Win Share Performance.}
+  \label{fig:WNBA_WS}
+\end{figure}
 
@@
 league average. If the value is less than 1, than the top 10 players in that statistic are paid the value as a factor below league average. The chart is 
-below:
-\vspace*{1cm}
-\begin{tabular}{*5l} 
+shown in \ref{tab:top10}.
+
+\begin{table}[tbp]
+    \caption{How Much Compared to Avg the Top 10 Players Are Paid For Advanced Stats.}
+    \label{tab:top10}
+\centering
+\begin{tabular}{crr}
 \toprule
-How Much Compared to Avg & the Top 10 Players Are Paid & For Advanced Stats \\\midrule
-
 Statistic &  WNBA  & NBA \\ 
-
+\midrule
  MP  &  1.445  & 5.353 \\
@@
  WS   &  1.530  & 5.446        \\
- \hline
+ \bottomrule
 \end{tabular}
+\end{table}
 
@@
 statistic in the WNBA and NBA datasets compared to player salaries. Since the base salary and proportional salary in the leagues are so different (as 
-shown in the aforementioned t-tests), it seemed most appropriate to simply see which statistics correlated the strongest to salary increases. Table 1 
+shown in the aforementioned t-tests), it seemed most appropriate to simply see which statistics correlated the strongest to salary increases. \ref{tab:main_data} 
 shows the selection of correlation coefficients for the thirteen matching statistics between the two league data sets.
@@
 \par
-The visual distribution of pay vs performance in the graphs is particularly interesting. While the correlation coefficient for both the WNBA and the NBA 
+The visual distribution of pay vs performance in figures \ref{fig:NBA_WS} and \ref{fig:WNBA_WS} is particularly interesting. While the correlation coefficient for both the WNBA and the NBA 
 when compared to Win Shares is relatively high, the NBA graph looks much more traditionally distributed: Examining it finds the data generally following 
-the trend line. The WNBA graph, however, looks much less traditional. It forms a sort of "U" shape, where players are almost all clustered into two salary 
+the trend line. The WNBA figure \ref{fig:WNBA_WS}, however, looks much less traditional. It forms a sort of "U" shape, where players are almost all clustered into two salary 
 buckets which are distributed similarly over the y axis. While there are more "higher performing" players in the higher pay bucket, the distribution is 
@@
 \par
-Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic impacts pay. As Table 2 shows, 
+Finally, some data manipulation was done to find a more simple statistic to try and determine how excellence in a statistic impacts pay. As \ref{tab:top10} shows, 
 there is a much clearer trend in performing in the top 10 in a statistic and a pay increase. As the WNBA pay spread is much smaller than the NBA’s it is 
@@
 \par
-Perhaps one of the most interesting conclusions is drawn when examining the graphs of NBA and WNBA pay distribution against Win Shares performance. 
+Perhaps one of the most interesting conclusions is drawn when examining figures \ref{fig:NBA_WS} and \ref{fig:WNBA_WS} of NBA and WNBA pay distribution against Win Shares performance. 
 Not only is there much less pay increase opportunity in the WNBA, but it seems as though almost all players are grouped into one of two pay rates, 
